### 2018-02-27 codes/R/slides/cohorts_nc_slides.Rnw
slides for NCDS conf

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=T, # REPLACE LATER
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@

<<READ, cache=FALSE >>=
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/1_load.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/2_efa.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_invariance.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_final.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores_compare.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/5_graphs.R")

@

<<PREAMBLE>>=
@

<<LOAD_DATA>>=
@

\documentclass[10pt,dvipsnames]{beamer}
\usepackage{mathpazo}  % serif math
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[absolute,overlay]{textpos}
\usepackage{booktabs}
\usepackage{bm}        % bold maths
\usepackage{dsfont}
\usepackage{ragged2e}
\usepackage{multirow}
\usepackage{natbib}
\usepackage[outdir=./]{epstopdf}
\usepackage{accents}  % accents in math mode
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{caption}
\usefonttheme[onlymath]{serif}
\usepackage{textgreek}

%tikz library
\usepackage{tikz}
\usetikzlibrary{intersections,calc,arrows,shapes}
\tikzset{>=latex}

% stata folder path
\newcommand{\statapics}{/Users/giacomomason/Dropbox/UCL/Whitehall/stata/w_graphs}

% additional symbols
\newcommand{\me}{\mathrm{e}}
\newcommand{\veps}{\varepsilon}
\newcommand{\E}{\mathbb{E}}
\newcommand{\1}{\mathds{1}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Corr}{\mathrm{Corr}}
\newcommand{\vect}{\mathrm{vec}}
\newcommand{\vecht}{\mathrm{vech}}
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}

% additional commands
\newcommand{\be}{\begin{eqnarray}}
\newcommand{\ee}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bnm}{\begin{enumerate}}
\newcommand{\enm}{\end{enumerate}}

% UCL beamer template
\definecolor{steelblue}{RGB}{70,130,180}
\definecolor{orangeucl}{RGB}{255,170,40}
\usetheme{ucl}
\setbeamertemplate{footline}[author title date]
\useoutertheme[small]{ucltitlebanner}
\setbeamertemplate{banner}[ucl][1]
\setbeamercolor{banner stripe}{bg=steelblue!70,fg=black}
\setbeamercolor{block body}{bg=steelblue!30,fg=black}
% override default square sizes to make smaller
\defbeamertemplate*{itemize item}{mysquare}{\tiny\raise0.5pt\hbox{$\blacksquare$}}
\defbeamertemplate*{itemize subitem}{mysquare}{\tiny\raise0.5pt\hbox{$\blacksquare$}}
\defbeamertemplate*{itemize subsubitem}{mysquare}{\tiny\raise0.5pt\hbox{$\blacksquare$}}
\setbeamertemplate{itemize item}[mysquare]               % override default square sizes
\setbeamertemplate{itemize subitem}[mysquare]
\setbeamertemplate{itemize subsubitem}[mysquare]

% define new columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\renewcommand{\arraystretch}{1.2}%

\setlength{\footnotesep}{0.4cm}

% reduced font size
\newcommand\Fontvi{\fontsize{8}{8}\selectfont} % reduce font size on slide
\newcommand\Fontvii{\fontsize{6}{6}\selectfont} % reduce font size on slide

% print TOC at every section break
\AtBeginSection[] { 
   \begin{frame}[noframenumbering]
   \frametitle{Outline}
   \tableofcontents[currentsection]
\end{frame}}

\title[Noncognitive skills in cohorts]{\bf{Comparability of child noncognitive skills across UK birth cohorts}}

\author[Mason]{Orazio Attanasio \and Richard Blundell \\ Gabriella Conti \and Giacomo Mason  }
\institute[UCL Economics]{UCL Department of Economics}

\date[HCRG]{Human Capital Reading Group \\ \today}

\begin{document}

\frame{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% INTRO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Motivation}
\framesubtitle{The role of noncognitive skills}
\bi
\item Traditional labour literature used to focus mainly on cognitive skills
    \bi 
    \item returns to ability in adulthood (labour market, health, \dots)
    \item process of child development
    \ei
\item Attention broadened to include \emph{non}-cognitive skills \citep{Cunha2007a,Cunha2010}
    \bi
    \item Socio-emotional, behavioural problems, character, personality traits, locus of control, grit, \dots
    \item Found to be very important in shaping adult outcomes \citep{Heckman2006,Carneiro2007,Heckman2012b}
    \item Malleable until later childhood
    \item Primary mechanism of transmission of effects from early life programmes to adult outcomes \citep{Heckman2013,Kautz2014a}
    \ei
\item Today, focus on behaviour problems and mental health
\ei
\end{frame}

\begin{frame}
\frametitle{Motivation}
\framesubtitle{Behavioural problems}

... something more on how behavioural problems are defined and fit within the literature on noncognitive skills

\end{frame}

\begin{frame}
\frametitle{Motivation}
\framesubtitle{Inequality in skills}

\bi
\item Socioeconomic gradients in behaviour problems are documented, and are persistent across childhood/adolescence \citep{Campbell1995,Segal2008}
\item Still not clear 
    \bi
    \item how these gradients change across time
    \item whether the returns to behavioural skills evolve as well
    \ei
\item \emph{Measurement} is a key (often overlooked) issue
    \bi
    \item Usually, behavioural constructs are measured via scales
    \item Most research uses (normalised) aggregate scores from scales \citep{Blanden2007}
    \item But does the same scale measure the same constructs in the same way:
      \bi 
      \item across different populations?
      \item in the same population, at different points in time?
      \ei
    \item And what if only different scales are available?
    \ei
\ei

\end{frame}

\begin{frame}
\frametitle{This study}

\bi
\item Focus on behavioural problems in two British cohorts
  \bi \item British Cohort Study (BCS, b. 1970)
  \item Millennium Cohort Study (MCS, b. 2000/1) \ei
\item Construct a common 11-item, 2-dimensional scale from different instruments
\item Investigate its psychometric properties (measurement invariance)
\item Use scores from the new scale to document changes in socioeconomic gradients between cohorts 30 years apart
\ei

\end{frame}

\begin{frame}
\frametitle{Outline}
  \tableofcontents
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% DATA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Data}

\begin{frame}
\frametitle{Data}
\framesubtitle{British Cohort Study}

\begin{columns}

\begin{column}{0.6\textwidth}
\bi
\item Data from 5 year follow-up, maternal questionnaire
\item Rutter A scale \citep{Rutter1970}
  \bi
  \item 19 items (behaviours)
  \item Options: \textit{Doesn't apply, Applies somewhat, Certainly applies}
  \item Usually conceptualised as measuring two dimensions
  \bi \item \emph{externalising} behaviour problems (anti-social) \item \emph{internalising} behaviour problems (neurotic) \ei
  \ei
  \item Additional questions from maternal module
\ei
\end{column}

\begin{column}{0.4\textwidth}
\includegraphics[width=\textwidth]{figs/bcsrut} 
\includegraphics[width=\textwidth]{figs/bcsrut2} 
\end{column}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Data}
\framesubtitle{Millennium Cohort Study}

\begin{columns}

\begin{column}{0.5\textwidth}

\bi
\item Data from 5 year follow-up (MCS3), maternal questionnaire
\item Complex survey design: need to rebalance data to make representative
\item Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994})
  \bi
  \item ``Updated" and expanded version of the Rutter scale 
  \item 25 items (behaviours)
  \item Options: \textit{Not true, Somewhat true, Certainly true}
  \item Designed with 5 subscales
  \item Broader subscales are sometimes justifiable \citep{Goodman2010}
  \ei
\ei

\end{column}

\begin{column}{0.5\textwidth}
\includegraphics[width=\textwidth]{figs/mcssdq} 
\end{column}

\end{columns}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ITEM CHOICE and MEAN TABLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Harmonisation}

\begin{frame}
\frametitle{Harmonisation}
\framesubtitle{Choice of items}

\bi
\item Aim: develop a new scale using only items that have similar wording across the two cohorts
\item Try to preserve three categories to maximise information
\item Combine items when needed (e.g. when SDQ items are positively coded)
\item Recode all items to higher categories for ``better" behaviours
\ei

\end{frame}

<<MEANTABLE>>=
# make table with means
@

\begin{frame}
\frametitle{Harmonisation}
\framesubtitle{Items}
\centering
\resizebox{!}{4cm}{
<<ITEM_TEXT, dependson="MEANTABLE", results='asis'>>=
require(xtable)

table <- meantab
table <- table[,!names(meantab) %in% c("cats", "fac", "bcs.num", "mcs.num")] # drop if unneeded

print(xtable(table,
                    align="L{0cm}C{.04\\textwidth}L{.4\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}L{.4\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 1
                    ), 
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
                                    paste(" & 
                                              \\multicolumn{4}{c}{\\textbf{BCS}} &
                                              \\multicolumn{4}{c}{\\textbf{MCS}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9} 
                                            Item &
                                              Text & Cert. Appl. & Smt. Appl. & Appl. &
                                              Text & Cert. True & Smt. True & True \\\\ \\midrule
                                          "
                                            , sep="")
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)

@
} %scalebox
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% EXPLORATORY FACTOR ANALYSIS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<BCSEFA>>=
@

<<BCSEFARES, dependson='BCSEFA'>>=
@

<<MCSEFA>>=
@

<<MCSEFARES, dependson='MCSEFA'>>=
@

\section{Exploratory Analysis}

\begin{frame}
\frametitle{EFA}
\framesubtitle{Number of factors}

\bi
\item Exploratory factor analysis on the two separate cohorts
\item Two-factor structure seems ok according to various approaches
\ei

\begin{center}
<<EFATAB, dependson='MCSEFA', results='asis'>>=
# combine results
efatab <- cbind(efatab.bcs,efatab.mcs)
colnames(efatab) <- c("BCS", "MCS")
rownames(efatab) <- c("Optimal Coordinates", "Acceleration Factor", "Parallel Analysis", "Kaiser", "VSS Compl. 1", "VSS Compl. 2")

print(xtable(efatab,
                                  # item num        Factor          Num cats        BCS text        BCS num           BCS Certain     BCS Somewhat    BCS Applies             MCS text          MCS num         
                    align="L{.4\\textwidth}C{.2\\textwidth}C{.2\\textwidth}",
                    digits = 0
                    ), 
             size="\\small",
             include.rownames=T,
             include.colnames=T,
             booktabs = T,
             sanitize.text.function = function(x){x},
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)
@
\end{center}
\end{frame}

\begin{frame}
\frametitle{EFA}
\framesubtitle{Factor structure}
\bi
\item In both cohorts, delivers a clear-cut split between two factors
\item Factors can be coherently interpreted as internalising and internalising behaviour problems
\ei 
\vspace{12pt}
\centering
\begin{tabular}{C{.05\textwidth}L{.3\textwidth}|C{.05\textwidth}L{.3\textwidth}}
  & \textbf{Externalising}  &    & \textbf{Internalising} \\
1 & Restless                & 7  & Worried                \\
2 & Squirmy/fidgety         & 8  & Afraid of new things   \\
3 & Fights/bullies children & 9  & Solitary               \\
4 & Cannot settle           & 10 & Unhappy/tearful        \\
5 & Temper tantrums         & 11 & Head/stomachaches \\
6 & Disobedient             &    &                       
\end{tabular}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Measurement Invariance}

\begin{frame}
\frametitle{Measurement Invariance}

\bi
\item Two constructs isolated in the new scale, separately by cohort
\item But are externalising and internalising problems measured in a comparable way \emph{across cohorts}?
  \bi
  \item Slightly different wording of question and options
  \item Different ordering in questionnaire
  \ei 
\vspace{12pt}\pause
\item \textbf{Measurement Invariance (MI)} analysis
  \bi
  \item Factor analysis -- but can also use IRT
  \item Pool the two cohorts together
  \item Start from minimally restrictive model, where parameters are free to vary across cohorts
  \item Place increasing restrictions on parameters across cohorts and test fit
  \ei 
\ei

\end{frame}

\begin{frame}
\frametitle{MI}\label{mimain}
\framesubtitle{Invariance steps}

\bi
\item Set up categorical Item Factor Analysis (IFA) model \hyperlink{miapp1}{\beamerbutton{Det.}}
\item Specify and compare model of different invariance levels

\bi
\item Approach in \citep{Wu2016} \hyperlink{miapp2}{\beamerbutton{Det.}}
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across cohorts)
  \ei
\item \textbf{Threshold and Loading invariance} model
  \bi
  \item Thresholds and loadings are restricted to be invariant across cohorts
  \item Means set to zero in both grups, variances set to 1 in reference cohort only
  \item Allows comparison of latent factor variances
  \ei
\item \textbf{Threshold, Loading, and Intercept invariance} model
  \bi
  \item Thresholds, loadings, intercepts restricted to be invariant across cohorts
  \item Mean and variances unrestricted in target cohort
  \item Allows comparison of latent factor means and variances
  \ei
\ei
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<FA_INVARIANCE, dependson='LOAD_DATA'>>=
# estimate models of different levels of invariance
@

<<FA_FIT, dependson='FA_INVARIANCE'>>=
# assemble fit indices into matrices
@

<<FA_FINAL, dependson='FA_INVARIANCE'>>=
# select final model set
@

<<FA_FINAL_PARS, dependson='FA_FINAL'>>=
# get parameters of final model
@

<<CLEANUP, dependson='FA_FINAL'>>=
@

<<FA_SCORES, dependson='FA_FINAL'>>=
# produce scores for different models
@

% # <<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
% # # assemble and compare scores across different models
% # @

<<FA_SCORES_SELECT, dependson='FA_FINAL'>>=
# select model to use for inequality analysis
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Fit comparison}
\begin{table}[ht]
\centering
\small
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[3]],afitab[[4]],afitab[[7]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}",
                    digits = 4
                    ),
             size="\\tiny",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                       \\textbf{Males only} \\\\ ",
                      "\\newline \\textbf{Females only} \\\\ \\midrule ",
                      "\\newline \\textbf{Males and females} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\end{table}
\end{frame}

\begin{frame}
\frametitle{MI}\label{mires}
\framesubtitle{Results}

\bi
\item Models with restricted intercepts do not pass MI test -- although the absolute fit is not terrible
  \bi
  \item $\Rightarrow$ cannot compare latent means according to \cite{Wu2016}
  \item but more traditional approaches with item anchoring would disagree \cite{Millsap2004}
  \ei
\item There are age differences between BCS and MCS due to survey design \hyperlink{agehist}{\beamerbutton{Graph}}
\item But restricting the age range to maximise overlap does not give significantly different results \hyperlink{fit4}{\beamerbutton{Graph}}
\item {\color{red} Can still use the model with non-invariant intercepts to score the latent factors and investigate inequality}
\ei

\end{frame}

\begin{frame}
\frametitle{Factor scores vs. Sum scores}
\begin{figure}
<<FACRAW, dev='tikz', fig.height=7, fig.width=7, out.width='.65\\linewidth'>>=
@
\end{figure}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% INEQUALITY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Inequality in behaviour problems}

\begin{frame}
\frametitle{Inequality}

\bi
\item Has inequality in behavioural problems increased between 1975 and 2005?
\item Investigate income gradient in externalising and internalising scores
  \bi
  \item Merge harmonised income data from CLOSER Work Package 2(a)
  \item Weekly family income when child is 10
  \item Would be more interesting with harmonised social class...
  \ei
\ei

\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient}
\begin{figure}
<<FACINEQ, dev='tikz', fig.height=6.7, fig.width=10, out.width='\\linewidth'>>=
@
\end{figure}
\end{frame}

<<LOESS_REAL, dependson="FA_SCORES_SELECT">>=
@

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_EXT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.ext
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_INT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.int
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation- and growth-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_REAL, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

scores2plot$incq_real <- factor(ntile(scores2plot$faminc_real, 5))

# get min, max
minmax.real <-aggregate(scores2plot[,c("faminc_real")],list(scores2plot$incq_real), 
            FUN = function(x) c(min = min(x, na.rm=T), max = max(x, na.rm=T)))
minmax.real <- round(minmax.real[,-1],2)*1000

r.ext.f.q <- lm(EXT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="F"))
r.int.f.q <- lm(INT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="F"))

r.ext.m.q <- lm(EXT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="M"))
r.int.m.q <- lm(INT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="M"))

stargazer(r.ext.m.q, r.int.m.q, r.ext.f.q, r.int.f.q,
          column.labels=c("Males","Females"), column.separate=c(2,2),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                paste0("Income Q2 [",toString(minmax.real[2,1]),"-",toString(minmax.real[2,2])," £/w]"),
                paste0("Income Q3 [",toString(minmax.real[3,1]),"-",toString(minmax.real[3,2])," £/w]"),
                paste0("Income Q4 [",toString(minmax.real[4,1]),"-",toString(minmax.real[4,2])," £/w]"),
                paste0("Income Q5 [",toString(minmax.real[5,1]),"-",toString(minmax.real[5,2])," £/w]"),
                             "MCS dummy",
                             "Income Q2 x MCS",
                             "Income Q3 x MCS",
                             "Income Q4 x MCS",
                             "Income Q5 x MCS"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

\begin{frame}
\frametitle{Inequality}

\bi
  \item The difference in scores between top and bottom quintiles has increased
  \item Scores at top quintiles have drifted upwards
\ei

\end{frame}

\begin{frame}
\frametitle{Returns}
\framesubtitle{Education returns to behaviour skills in BCS}
\centering
\resizebox{!}{4cm}{%
<<FA_RETURNS, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

# get BCS data and make dummy for higher education
scoresBCS <- scores2plot[scores2plot$cohort=="BCS",]
scoresBCS$highered <- as.numeric(scoresBCS$hinvq00>3)
scoresBCS[is.na(scoresBCS$hinvq00), "highered"] <- NA

# females
r.nvq.f <- list()
r.nvq.f[[1]] <- lm(highered ~ EXT + INT, data=subset(scoresBCS, sex=="F"))
r.nvq.f[[2]] <- lm(highered ~ EXT + INT + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="F"))
r.nvq.f[[3]] <- lm(highered ~ EXT + INT + factor(incq), data=subset(scoresBCS, sex=="F"))
r.nvq.f[[4]] <- lm(highered ~ EXT + INT + ysch_moth + ysch_fath, data=subset(scoresBCS, sex=="F"))
r.nvq.f[[5]] <- lm(highered ~ EXT + INT + factor(incq) + ysch_moth + ysch_fath + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="F"))

# females
r.nvq.m <- list()
r.nvq.m[[1]] <- lm(highered ~ EXT + INT, data=subset(scoresBCS, sex=="M"))
r.nvq.m[[2]] <- lm(highered ~ EXT + INT + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="M"))
r.nvq.m[[3]] <- lm(highered ~ EXT + INT + factor(incq), data=subset(scoresBCS, sex=="M"))
r.nvq.m[[4]] <- lm(highered ~ EXT + INT + ysch_moth + ysch_fath, data=subset(scoresBCS, sex=="M"))
r.nvq.m[[5]] <- lm(highered ~ EXT + INT + factor(incq) + ysch_moth + ysch_fath + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="M"))

stargazer(r.nvq.m[[1]], r.nvq.m[[2]], r.nvq.m[[3]], r.nvq.m[[4]], r.nvq.m[[5]],
          r.nvq.f[[1]], r.nvq.f[[2]], r.nvq.f[[3]], r.nvq.f[[4]], r.nvq.f[[5]],
          dep.var.caption = "Completed Higher Ed. at 30",
          column.labels=c("Males","Females"), column.separate=c(5,5),
          omit=c("Constant"),
          covariate.labels=c("Externalising skill at 5",
                             "Internalising skill at 5",
                             "EPVT Z at 5",
                             "Human figure drawing Z at 5",
                             "Copy design Z at 5",
                             "Family inc. at 10 Q2",
                             "Family inc. at 10 Q3",
                             "Family inc. at 10 Q4",
                             "Family inc. at 10 Q5",
                             "Mother's years of schooling",
                             "Father's years of schooling"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[noframenumbering]
\centering
{\Large Appendix}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp1}
\framesubtitle{IFA model}
Categorical Item Factor Analysis (IFA) model \hyperlink{mimain}{\beamerreturnbutton{Back}}

\bi
\item Threshold model \citep{Muthen1984}
\item Latent propensity to endorse each item $i$ for child $j$ in cohort $c$
\begin{align}\label{eq:propens}
\begin{split}
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{EXT}_{jc} + u_{ijc} \qquad \text{for} \; i=1,\dots, 6 \\
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{INT}_{jc} + u_{ijc} \qquad \text{for} \; i=7,\dots, 11
\end{split}
\end{align}

\item Mean/covariance structure
$$\bm{\mu}_c=\bm{\nu}_c + \bm{\Lambda}_c\bm{\kappa}_c \qquad \text{and} \qquad \Sigma_c = \bm{\Lambda}_c \bm{\Phi}_c \bm{\Lambda}_c^\prime + \bm{\Psi}_c.$$
with $$(\theta^{EXT}_{jc}, \theta^{INT}_{jc}) = \bm{\Theta}_{jc} \sim N(\bm{\kappa}_c, \bm{\Phi}_c) \qquad \text{and} \qquad \bm{u}_{jc} \sim N(\bm{0}, \bm{\Psi}_c)$$

\item Discreteness
$$X_{ijc} = s \qquad \text{if} \; \tau_{s,ic} \leq X^*_{ijc} < \tau_{s+1,ic} \qquad \text{for} \; s=0,1,2$$
with $\tau_{0,ic}=-\infty$ and $\tau_{3,ic}=+\infty$.

\ei
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Key Parameters for Invariance}

\bi
\item Loadings $\lambda_{ci}$
\item Thresholds $\tau_{1,ci}$, $\tau_{2,ci}$
\item Intercepts $\nu_{ci}$
\item Unique variances $\Var(u_{ci})$
\item Latent mean vector $\bm{\kappa}_c$
\item Latent covariance matrix $\bm{\Phi}_c$
\ei

\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp2}
\framesubtitle{Invariance steps \hyperlink{mimain}{\beamerreturnbutton{Back}}}
\Fontvi
\bi
\item \textbf{Configural} model
$$\text{diag}(\bm{\Phi}) = \bm{I}, \quad \bm{\kappa}=\bm{0}, \quad \bm{\nu}=\bm{0}, \quad \text{and} \; \text{diag}(\bm{\Psi})=\bm{I}.$$

\item \textbf{Threshold invariance} model
  \bi
  \item $\tau_{1,Ai} = \tau_{1,Bi}$ for $i=1,\dots,11$, $\tau_{2,Ai} = \tau_{2,Bi}$ for $i=\{1,2,3,4,7,8,9,10\}$
  \item For all groups: $\text{diag}(\bm{\Phi}) = \bm{I}, \quad \bm{\kappa}=\bm{0}$
  \item For the reference group: $\bm{\nu}_A=\bm{0}, \quad \; \text{diag}(\bm{\Sigma}_A)=\bm{I}$
  \ei
\item \textbf{Threshold and Loading invariance} model
  \bi
  \item $\tau_{1,Ai} = \tau_{1,Bi}$ for $i=1,\dots,11$, $\tau_{2,Ai} = \tau_{2,Bi}$ for $i=\{1,2,3,4,7,8,9,10\}$
  \item $\lambda_{Ai} = \lambda_{Bi}$ for $i=1,\dots,11$
  \item For all groups: $\bm{\kappa}=\bm{0}$
  \item For the reference group: $\bm{\nu}_A=\bm{0}, \quad \; \text{diag}(\bm{\Sigma}_A)=\bm{I}, \quad \text{diag}(\bm{\Phi}_A) = \bm{I}$
  \ei
\item \textbf{Threshold, Loading, and Intercept invariance} model
  \bi
  \item $\tau_{1,Ai} = \tau_{1,Bi}$ for $i=1,\dots,11$, $\tau_{2,Ai} = \tau_{2,Bi}$ for $i=\{1,2,3,4,7,8,9,10\}$
  \item $\lambda_{Ai} = \lambda_{Bi}$ for $i=1,\dots,11$
  \item For all groups: $\bm{\nu}=\bm{0}$
  \item For the reference group: $\bm{\kappa}_A=\bm{0}, \quad \; \text{diag}(\bm{\Sigma}_A)=\bm{I}, \quad \text{diag}(\bm{\Phi}_A) = \bm{I}$
  \ei
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ROBUSTNESS APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[noframenumbering]
\frametitle{Age distribution \hyperlink{mires}{\beamerreturnbutton{Back}}}\label{agehist}
\begin{figure}
\begin{center}
<<AGEHIST, dev='tikz', fig.height=6, fig.width=6, out.width='.6\\linewidth', cache=TRUE>>=
@
\end{center}
\caption{Distribution of age in the BCS and MCS sample}
\end{figure}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Fit comparison -- Robustness}
List of estimated models:
\bnm
\item no gender split, no age adjustment (2 groups, BCS MCS)
\item no gender split, age adjustment (2 groups, BCS MCS)
\item males only, no age adjustment (2 groups, BCS MCS)
\item females only, no age adjustment (2 groups, BCS MCS)
\item males only, age adjustment (2 groups, BCS MCS)
\item females only, age adjustment (2 groups, BCS MCS)
\item separate gender groups, no age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, overlapping ages, no age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, age adjustment constrained to be the same across ages (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item MIMIC model with age as regressor (4 groups, BCS.M BCS.F MCS.M MCS.F)
\enm

\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Fit comparison -- 2 group models, age unadjusted}

\begin{table}[ht]
\centering
\small
<<FA_FIT_RES2a, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[1]],afitab[[3]],afitab[[4]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}",
                    digits = 4
                    ),
             size="\\tiny",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                       \\textbf{(1) No gender split, not age adjusted} \\\\",
                      "\\newline \\textbf{(3) Males only, not age adjusted} \\\\ \\midrule ",
                      "\\newline \\textbf{(4) Females only, not age adjusted} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\end{table}

\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Fit comparison -- 2 group models, age adjusted}
\begin{table}[ht]
\centering
\small
<<FA_FIT_RES2b, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[2]],afitab[[5]],afitab[[6]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}",
                    digits = 4
                    ),
             size="\\tiny",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                       \\textbf{(2) No gender split, age adjusted} \\\\ ",
                      "\\newline \\textbf{(5) Males only, age adjusted} \\\\ \\midrule ",
                      "\\newline \\textbf{(6) Females only, age adjusted} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\end{table}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{fit4}
\framesubtitle{Fit comparison -- 4 group models \hyperlink{mires}{\beamerreturnbutton{Back}}}

\begin{table}[ht]
\centering
\small
<<FA_FIT_RES4, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[7]],afitab[[8]],afitab[[9]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}",
                    digits = 4
                    ),
             size="\\tiny",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                      \\newline \\textbf{(7) not age adjusted} \\\\",
                      "\\newline \\textbf{(8) age adjusted} \\\\\\midrule ",
                      "\\newline \\textbf{(9) 59-62m range} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\end{table}

\end{frame}

<<LOESS_INFL, dependson="FA_SCORES_SELECT">>=
@

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_EXT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.ext
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_INT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.int
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_INFL, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

scores2plot$incq_infl <- factor(ntile(scores2plot$faminc_infl, 5))

r.ext.f.c <- lm(EXT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="F"))
r.ext.f.q <- lm(EXT ~ incq_infl*cohort, data=subset(scores2plot, sex=="F"))
r.int.f.c <- lm(INT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="F"))
r.int.f.q <- lm(INT ~ incq_infl*cohort, data=subset(scores2plot, sex=="F"))

r.ext.m.c <- lm(EXT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="M"))
r.ext.m.q <- lm(EXT ~ incq_infl*cohort, data=subset(scores2plot, sex=="M"))
r.int.m.c <- lm(INT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="M"))
r.int.m.q <- lm(INT ~ incq_infl*cohort, data=subset(scores2plot, sex=="M"))

stargazer(r.ext.m.q, r.int.m.q, r.ext.f.q, r.int.f.q,
          column.labels=c("Males","Females"), column.separate=c(2,2),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                             "Income Q2",
                             "Income Q3",
                             "Income Q4",
                             "Income Q5",
                             "MCS dummy",
                             "Income Q2 x MCS",
                             "Income Q3 x MCS",
                             "Income Q4 x MCS",
                             "Income Q5 x MCS"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

\begin{frame}[noframenumbering]
\bibliographystyle{chicago}
\def\bibfont{\tiny}
\bibliography{/Users/giacomomason/Documents/BibTex/Papers-cohorts_nc}
\end{frame}

\end{document}