### 2018-05-19 codes/R/slides/cohorts_tapes.Rnw
removed some variables

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=T, # REPLACE LATER
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@

<<READ, cache=FALSE >>=
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/1_load.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/2_efa.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_invariance.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_final.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores_compare.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/5_graphs.R")

@

<<PREAMBLE>>=
@

<<LOAD_DATA>>=
@

\documentclass[10pt,dvipsnames]{beamer}
\usepackage{mathpazo}  % serif math
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[absolute,overlay]{textpos}
\usepackage{booktabs}
\usepackage{bm}        % bold maths
\usepackage{dsfont}
\usepackage{ragged2e}
\usepackage{multirow}
\usepackage{natbib}
\usepackage[outdir=./]{epstopdf}
\usepackage{accents}  % accents in math mode
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{caption}
\usefonttheme[onlymath]{serif}
\usepackage{textgreek}

%tikz library
\usepackage{tikz}
\usetikzlibrary{intersections,calc,arrows,shapes}
\tikzset{>=latex}

% stata folder path
\newcommand{\statapics}{/Users/giacomomason/Dropbox/UCL/Whitehall/stata/w_graphs}

% additional symbols
\newcommand{\me}{\mathrm{e}}
\newcommand{\veps}{\varepsilon}
\newcommand{\E}{\mathbb{E}}
\newcommand{\1}{\mathds{1}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Corr}{\mathrm{Corr}}
\newcommand{\vect}{\mathrm{vec}}
\newcommand{\vecht}{\mathrm{vech}}
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}

% additional commands
\newcommand{\be}{\begin{eqnarray}}
\newcommand{\ee}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bnm}{\begin{enumerate}}
\newcommand{\enm}{\end{enumerate}}

% UCL beamer template
\definecolor{deepskyblue}{RGB}{0,133,178}
\definecolor{orangeucl}{RGB}{255,170,40}
\usetheme{ucl}
\setbeamertemplate{footline}[author title date]
\useoutertheme[small]{ucltitlebanner}
\setbeamertemplate{banner}[ucl][1]
\setbeamercolor{banner stripe}{bg=deepskyblue!70,fg=black}
\setbeamercolor{block body}{bg=steelblue!30,fg=black}
% override default square sizes to make smaller
\defbeamertemplate*{itemize item}{mysquare}{\tiny\raise0.5pt\hbox{$\blacksquare$}}
\defbeamertemplate*{itemize subitem}{mysquare}{\tiny\raise0.5pt\hbox{$\blacksquare$}}
\defbeamertemplate*{itemize subsubitem}{mysquare}{\tiny\raise0.5pt\hbox{$\blacksquare$}}
\setbeamertemplate{itemize item}[mysquare]               % override default square sizes
\setbeamertemplate{itemize subitem}[mysquare]
\setbeamertemplate{itemize subsubitem}[mysquare]

% define new columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\renewcommand{\arraystretch}{1.2}%

\setlength{\footnotesep}{0.4cm}

% reduced font size
\newcommand\Fontvi{\fontsize{8}{8}\selectfont} % reduce font size on slide
\newcommand\Fontvii{\fontsize{6}{6}\selectfont} % reduce font size on slide

% print TOC at every section break
\AtBeginSection[] { 
   \begin{frame}[noframenumbering]
   \frametitle{Outline}
   \tableofcontents[currentsection]
\end{frame}}

\title[Noncognitive skills in cohorts]{\bf{Inequality in non-cognitive skills: \\ a cross-cohort comparison}}

\author[Mason]{Orazio Attanasio \and Richard Blundell \\ Gabriella Conti \and Giacomo Mason  }
\institute[UCL Economics]{UCL Department of Economics}

\date[NCDS]{NCDS 60 Years of Our Lives \\ \today}

\begin{document}

\frame{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% INTRO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Motivation}
\framesubtitle{Inequality in skills}

\bi
\item Non-cognitive (socio-emotional, behavioural) skills developed during childhood are important for lifetime outcomes
\item Socioeconomic gradients in these skills are persistent 
\pause 
\item \textbf{Q1}: Can we derive \textbf{comparable measures} across populations?
    \bi
    \item Usually, behavioural constructs are measured via scales
    \item But does the same scale measure the same constructs in the same way:
      \bi 
      \item across different populations?
      \item in the same population, at different points in time?
      \ei
    \item And what if only different scales are available?
    \ei
\pause
\item \textbf{Q2}: Can we use these measures to document \textbf{changes in socioeconomic inequality} in skills across populations?
\ei

\end{frame}

\begin{frame}
\frametitle{This study}

\bi
\item Focus on behavioural problems in two British cohorts
  \bi \item British Cohort Study (BCS, b. 1970)
  \item Millennium Cohort Study (MCS, b. 2000/1) \ei
\ei
\vspace{6pt} \pause
\bnm
\item Construct a common 11-item scale of non-cognitive skills, starting from different instruments
\item Investigate its psychometric properties
\item Use scores from the new scale to document changes in socioeconomic gradients between cohorts 30 years apart
\enm
\vspace{6pt} \pause
\bi
\item \textbf{Key findings}:
  \bi
  \item Inequality in non-cognitive skills has increased in 1975-2005
  \item Sharpest change for boys rather than girls
  \item Mostly driven by improvements at the top of the distribution
  \ei
\ei

\end{frame}

\begin{frame}
\frametitle{Outline}
  \tableofcontents
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% DATA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Harmonised scale}

\subsection{Data}

\begin{frame}
\frametitle{Data}
\framesubtitle{British Cohort Study}

\begin{columns}

\begin{column}{0.6\textwidth}
\bi
\item Data from 5 year follow-up (England only), maternal questionnaire
\item Rutter A scale \citep{Rutter1970}
  \bi
  \item 19 items (behaviours)
  \item Options: \textit{Doesn't apply, Applies somewhat, Certainly applies}
  \item Usually conceptualised as measuring two dimensions
  \bi \item \emph{externalising} behaviour problems (anti-social) \item \emph{internalising} behaviour problems (neurotic) \ei
  \ei
  \item Additional questions from maternal module
\ei
\end{column}

\begin{column}{0.4\textwidth}
\includegraphics[width=\textwidth]{figs/bcsrut} 
\includegraphics[width=\textwidth]{figs/bcsrut2} 
\end{column}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Data}
\framesubtitle{Millennium Cohort Study}

\begin{columns}

\begin{column}{0.5\textwidth}

\bi
\item Data from 5 year follow-up (England only), maternal questionnaire
  \bi \item Complex survey design: rebalance sample to make representative \ei
\item Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994})
  \bi
  \item ``Updated" and expanded version of the Rutter scale 
  \item 25 items (behaviours)
  \item Options: \textit{Not true, Somewhat true, Certainly true}
  \item Designed with 5 subscales
  \item Broader subscales are sometimes justifiable \citep{Goodman2010}
  \ei
\ei

\end{column}

\begin{column}{0.5\textwidth}
\includegraphics[width=\textwidth]{figs/mcssdq} 
\end{column}

\end{columns}

\end{frame}

\subsection{Common items}

<<MEANTABLE>>=
# make table with means
@

\begin{frame}
\frametitle{Harmonisation}
\framesubtitle{Choice of common items}
\centering
\resizebox{!}{4cm}{
<<ITEM_TEXT, dependson="MEANTABLE", results='asis'>>=
require(xtable)

table <- meantab
table <- table[,!names(meantab) %in% c("cats", "fac", "bcs.num", "mcs.num")] # drop if unneeded

print(xtable(table,
                    align="L{0cm}C{.04\\textwidth}L{.4\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}L{.4\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 1
                    ), 
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
                                    paste(" & 
                                              \\multicolumn{4}{c}{\\textbf{BCS}} &
                                              \\multicolumn{4}{c}{\\textbf{MCS}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9} 
                                            Item &
                                              Text & Cert. Appl. & Smt. Appl. & Appl. &
                                              Text & Cert. True & Smt. True & True \\\\ \\midrule
                                          "
                                            , sep="")
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)

@
} %scalebox
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% EXPLORATORY FACTOR ANALYSIS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<BCSEFA>>=
@

<<BCSEFARES, dependson='BCSEFA'>>=
@

<<MCSEFA>>=
@

<<MCSEFARES, dependson='MCSEFA'>>=
@

\subsection{Exploratory Analysis}

\begin{frame}
\frametitle{EFA}\label{efa}
\framesubtitle{Factor structure}

\bi
\item Different methods for assessing number of factors give varied results \hyperlink{numfac}{\beamerbutton{Det.}}
\item We choose two factors to maximise interpretability
\item In both cohorts, EFA delivers a clear-cut split between two factors
\item Factors can be coherently interpreted as internalising and internalising skills
\ei 
\vspace{12pt}
\centering
\begin{tabular}{C{.05\textwidth}L{.3\textwidth}|C{.05\textwidth}L{.3\textwidth}}
  & \textbf{Externalising}  &    & \textbf{Internalising} \\
1 & Restless                & 7  & Worried                \\
2 & Squirmy/fidgety         & 8  & Afraid of new things   \\
3 & Fights/bullies children & 9  & Solitary               \\
4 & Cannot settle           & 10 & Unhappy/tearful        \\
5 & Temper tantrums         & 11 & Head/stomachaches \\
6 & Disobedient             &    &                       
\end{tabular}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Measurement Invariance}

\begin{frame}
\frametitle{Measurement Invariance}\label{mimain}
\framesubtitle{Model setup}

\bi
\item Two constructs isolated in the new scale, separately by cohort
\item But are externalising and internalising problems measured in a comparable way \emph{across cohorts}?
\vspace{12pt}\pause
\item \textbf{Measurement Invariance (MI)} analysis
  \bi
  \item Formal test of the invariance hypothesis
  \item Set up categorical Item Factor Analysis (IFA) model \hyperlink{miapp1}{\beamerbutton{Det.}}
  \item Pool the two cohorts together, and estimate models of different invariance level -- following \cite{Wu2016}:\hyperlink{miapp2}{\beamerbutton{Det.}}
    \bnm
    \item Start from minimally restrictive model, where parameters are free to vary
    \item Restrict threshold and loading parameters to be equal across cohorts
    \item Restrict intercept parameters as well
    \enm
  \item Test relative fit of restricted models compared to initial model
  \ei 
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<FA_INVARIANCE>>=
# estimate models of different levels of invariance
@

<<FA_FIT, dependson='FA_INVARIANCE'>>=
# assemble fit indices into matrices
@

<<FA_FINAL, dependson='FA_INVARIANCE'>>=
# select final model set
@

<<FA_FINAL_PARS, dependson='FA_FINAL'>>=
# get parameters of final model
@

<<CLEANUP, dependson='FA_FINAL'>>=
@

<<FA_SCORES, dependson='FA_FINAL'>>=
# produce scores for different models
@

<<FA_SCORES_MERGE, dependson='LOAD_DATA'>>=
# merge scores with other data
@

<<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
# assemble and compare scores across different models
@

<<FA_SCORES_SELECT, dependson='FA_SCORES_COMPARE'>>=
# select model to use for inequality analysis
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Measurement Invariance}\label{mires}
\framesubtitle{Results}

\bi
\item Restricting thresholds and loadings provides a good fit
  \bi \item We can compare the scale (variance) of the latent factors \ei
\item Models with restricted intercepts do \emph{not} pass MI test
  \bi
  \item Unclear whether mean comparisons can be done
  \item \cite{Wu2016} vs \cite{Millsap2004}
  \ei
% \item Might \emph{age differences} in the sample be playing a role?
%   \bi
%   \item BCS and MCS samples differ due to survey design \hyperlink{agehist}{\beamerbutton{Graph}}
%   \item But restricting the age range to maximise overlap does not give significantly different results \hyperlink{fit4}{\beamerbutton{Det.}}
%   \ei
  \pause
\item \textbf{Q1}: Can we derive comparable measures of noncognitive skills across cohorts?
\item Yes, we can use the model with non-invariant intercepts to score the latent factors (externalising and internalising skills)
\ei

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% INEQUALITY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Inequality in behaviour problems}

\begin{frame}
\frametitle{Inequality}

\bi
\item \textbf{Q2}: Has socioeconomic inequality in behavioural problems increased between 1975 and 2005?
\item Investigate income gradient in externalising and internalising scores
  \bi
  \item Merge harmonised income data from CLOSER Work Package 2(a)
  \item Weekly family income when child is 10 (not contemporaneous)
  \item Working on access to harmonised social class data
  \ei
\ei

\end{frame}

\begin{frame}
\frametitle{Inequality}\label{meanci}
\framesubtitle{Visualising the gradient \hyperlink{loessext}{\beamerbutton{EXT}}\hyperlink{loessint}{\beamerbutton{INT}}}
\begin{figure}
<<FACINEQ, dev='tikz', fig.height=6.7, fig.width=10, out.width='\\linewidth'>>=
@
\end{figure}
\end{frame}

% \begin{frame}
% \frametitle{Inequality}
% \framesubtitle{Quantifying the gradient -- inflation- and growth-adjusted income}
% \centering
% \resizebox{!}{4cm}{%
% <<FA_REGS_REAL, dependson='FA_SCORES_SELECT', results='asis'>>=
% # regress scores on income
% library(dplyr)
% library(stargazer)
% 
% scores2plot$incq_real <- factor(ntile(scores2plot$faminc_real, 5))
% 
% # get min, max
% minmax.real <-aggregate(scores2plot[,c("faminc_real")],list(scores2plot$incq_real), 
%             FUN = function(x) c(min = min(x, na.rm=T), max = max(x, na.rm=T)))
% minmax.real <- round(minmax.real[,-1],2)*1000
% 
% r.ext.f.q <- lm(EXT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="F"))
% r.int.f.q <- lm(INT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="F"))
% 
% r.ext.m.q <- lm(EXT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="M"))
% r.int.m.q <- lm(INT ~ as.factor(incq_real)*cohort, data=subset(scores2plot, sex=="M"))
% 
% stargazer(r.ext.m.q, r.int.m.q, r.ext.f.q, r.int.f.q,
%           column.labels=c("Males","Females"), column.separate=c(2,2),
%           dep.var.caption = "", multicolumn = FALSE,
%           omit=c("Constant"),
%           covariate.labels=c(
%                 paste0("Income Q2 [",toString(minmax.real[2,1]),"-",toString(minmax.real[2,2])," £/w]"),
%                 paste0("Income Q3 [",toString(minmax.real[3,1]),"-",toString(minmax.real[3,2])," £/w]"),
%                 paste0("Income Q4 [",toString(minmax.real[4,1]),"-",toString(minmax.real[4,2])," £/w]"),
%                 paste0("Income Q5 [",toString(minmax.real[5,1]),"+ £/w]"),
%                              "MCS dummy",
%                              "Income Q2 x MCS",
%                              "Income Q3 x MCS",
%                              "Income Q4 x MCS",
%                              "Income Q5 x MCS"
%                              ),
%           no.space=TRUE, 
%           omit.stat= c("rsq", "f", "ser", "aic", "ll"),
%           float=FALSE    # needs begin table outside it
% )
% @
% }
% \end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient}
\centering
\resizebox{12cm}{!}{%
<<FA_REGS_REAL_MF, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

r.ext.m.bcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="BCS"))
r.ext.m.mcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="MCS"))
r.int.m.bcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="BCS"))
r.int.m.mcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="MCS"))

r.ext.f.bcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="BCS"))
r.ext.f.mcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="MCS"))
r.int.f.bcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="BCS"))
r.int.f.mcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="MCS"))

stargazer(r.ext.m.bcs, r.ext.m.mcs , r.int.m.bcs, r.int.m.mcs, r.ext.f.bcs, r.ext.f.mcs , r.int.f.bcs, r.int.f.mcs,
          column.labels=c("BCS M","MCS M","BCS M","MCS M","BCS F","MCS F","BCS F","MCS F"),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                             "Income Q2",
                             "Income Q3",
                             "Income Q4",
                             "Income Q5"
                             ),
          no.space=TRUE,
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

% 
% \begin{frame}
% \frametitle{Inequality}
% \framesubtitle{Quantifying the gradient -- males}
% \centering
% \resizebox{12cm}{4cm}{%
% <<FA_REGS_REAL_M, dependson='FA_SCORES', results='asis'>>=
% # regress scores on income
% library(dplyr)
% library(stargazer)
% 
% 
% r.ext.m.bcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="BCS"))
% r.ext2.m.bcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="BCS"))
% r.ext.m.mcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="MCS"))
% r.ext2.m.mcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="MCS"))
% r.int.m.bcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="BCS"))
% r.int2.m.bcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="BCS"))
% r.int.m.mcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="MCS"))
% r.int2.m.mcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="MCS"))
% 
% stargazer(r.ext.m.bcs, r.ext2.m.bcs, r.int.m.bcs, r.int2.m.bcs, r.ext.m.mcs, r.ext2.m.mcs, r.int.m.mcs, r.int2.m.mcs,
%           column.labels=c("BCS","MCS"), column.separate=c(4,4),
%           dep.var.caption = "", multicolumn = FALSE,
%           omit=c("Constant"),
%           covariate.labels=c(
%                              "Income Q2",
%                              "Income Q3",
%                              "Income Q4",
%                              "Income Q5",
%                              "Birthweight",
%                              "Mat. Smoking in preg.",
%                              "Father years of schooling",
%                              "Mother years of schooling"
%                              ),
%           no.space=TRUE, 
%           omit.stat= c("rsq", "f", "ser", "aic", "ll"),
%           float=FALSE    # needs begin table outside it
% )
% @
% }
% \end{frame}
% 
% 
% 
% \begin{frame}
% \frametitle{Inequality}
% \framesubtitle{Quantifying the gradient -- females}
% \centering
% \resizebox{12cm}{4cm}{%
% <<FA_REGS_REAL_F, dependson='FA_SCORES', results='asis'>>=
% # regress scores on income
% library(dplyr)
% library(stargazer)
% 
% 
% r.ext.f.bcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="BCS"))
% r.ext2.f.bcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="BCS"))
% r.ext.f.mcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="MCS"))
% r.ext2.f.mcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="MCS"))
% r.int.f.bcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="BCS"))
% r.int2.f.bcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="BCS"))
% r.int.f.mcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="MCS"))
% r.int2.f.mcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="MCS"))
% 
% 
% 
% stargazer(r.ext.f.bcs, r.ext2.f.bcs, r.int.f.bcs, r.int2.f.bcs, r.ext.f.mcs, r.ext2.f.mcs, r.int.f.mcs, r.int2.f.mcs,
%           column.labels=c("BCS","MCS"), column.separate=c(4,4),
%           dep.var.caption = "", multicolumn = FALSE,
%           omit=c("Constant"),
%           covariate.labels=c(
%                              "Income Q2",
%                              "Income Q3",
%                              "Income Q4",
%                              "Income Q5",
%                              "Birthweight",
%                              "Mat. Smoking in preg.",
%                              "Father years of schooling",
%                              "Mother years of schooling"
%                              ),
%           no.space=TRUE, 
%           omit.stat= c("rsq", "f", "ser", "aic", "ll"),
%           float=FALSE    # needs begin table outside it
% )
% @
% }
% \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% CONCLUSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions and next steps}

\begin{frame}
\frametitle{Conclusions and next steps}

\bi
\item We derived a comparable scale of non-cognitive skills, based on maternal reports of behaviour problems
\item We have shown that inequality in these skills at five years of age has increased between children born in 1970 and 2000
  \bi
  \item Profile has become steeper (particularly for males)
  \item Scores at upper quantiles have drifted upwards
  \ei
\pause
\vspace{12pt}
\item Next steps
  \bi
  \item Harmonise more cohorts (NCDS) and ages (10)
  \item Broaden analysis to the \emph{returns} to non-cognitive skills -- are these remunerated differently on the labour market for people born in different years?
  \ei
\ei

\end{frame}

\begin{frame}[noframenumbering]
\centering
{\Large Thank you!}
\end{frame}

% \begin{frame}
% \frametitle{Returns}
% \framesubtitle{Education returns to behaviour skills in BCS}
% \centering
% \resizebox{!}{4cm}{%
% <<FA_RETURNS, dependson='FA_SCORES_SELECT', results='asis'>>=
% # regress scores on income
% library(dplyr)
% library(stargazer)
% 
% # get BCS data and make dummy for higher education
% scoresBCS <- scores2plot[scores2plot$cohort=="BCS",]
% scoresBCS$highered <- as.numeric(scoresBCS$hinvq00>3)
% scoresBCS[is.na(scoresBCS$hinvq00), "highered"] <- NA
% 
% # females
% r.nvq.f <- list()
% r.nvq.f[[1]] <- lm(highered ~ EXT + INT, data=subset(scoresBCS, sex=="F"))
% r.nvq.f[[2]] <- lm(highered ~ EXT + INT + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="F"))
% r.nvq.f[[3]] <- lm(highered ~ EXT + INT + factor(incq), data=subset(scoresBCS, sex=="F"))
% r.nvq.f[[4]] <- lm(highered ~ EXT + INT + ysch_moth + ysch_fath, data=subset(scoresBCS, sex=="F"))
% r.nvq.f[[5]] <- lm(highered ~ EXT + INT + factor(incq) + ysch_moth + ysch_fath + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="F"))
% 
% # females
% r.nvq.m <- list()
% r.nvq.m[[1]] <- lm(highered ~ EXT + INT, data=subset(scoresBCS, sex=="M"))
% r.nvq.m[[2]] <- lm(highered ~ EXT + INT + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="M"))
% r.nvq.m[[3]] <- lm(highered ~ EXT + INT + factor(incq), data=subset(scoresBCS, sex=="M"))
% r.nvq.m[[4]] <- lm(highered ~ EXT + INT + ysch_moth + ysch_fath, data=subset(scoresBCS, sex=="M"))
% r.nvq.m[[5]] <- lm(highered ~ EXT + INT + factor(incq) + ysch_moth + ysch_fath + cog1 + cog2 + cog3, data=subset(scoresBCS, sex=="M"))
% 
% stargazer(r.nvq.m[[1]], r.nvq.m[[2]], r.nvq.m[[3]], r.nvq.m[[4]], r.nvq.m[[5]],
%           r.nvq.f[[1]], r.nvq.f[[2]], r.nvq.f[[3]], r.nvq.f[[4]], r.nvq.f[[5]],
%           dep.var.caption = "Completed Higher Ed. at 30",
%           column.labels=c("Males","Females"), column.separate=c(5,5),
%           omit=c("Constant"),
%           covariate.labels=c("Externalising skill at 5",
%                              "Internalising skill at 5",
%                              "EPVT Z at 5",
%                              "Human figure drawing Z at 5",
%                              "Copy design Z at 5",
%                              "Family inc. at 10 Q2",
%                              "Family inc. at 10 Q3",
%                              "Family inc. at 10 Q4",
%                              "Family inc. at 10 Q5",
%                              "Mother's years of schooling",
%                              "Father's years of schooling"
%                              ),
%           no.space=TRUE, 
%           omit.stat= c("rsq", "f", "ser", "aic", "ll"),
%           float=FALSE    # needs begin table outside it
% )
% @
% }
% \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[noframenumbering]
\centering
{\Large Appendix}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{EFA}\label{numfac}
\framesubtitle{Number of factors \hyperlink{efa}{\beamerreturnbutton{Back}}}
\begin{center}
<<EFATAB, dependson='MCSEFA', results='asis'>>=
# combine results
efatab <- cbind(efatab.bcs,efatab.mcs)
colnames(efatab) <- c("BCS All", "BCS M", "BCS F", "MCS All", "MCS M", "MCS F")

print(xtable(efatab,
                    align="L{.25\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 0
                    ), 
             size="\\small",
             include.rownames=T,
             include.colnames=T,
             booktabs = T,
             sanitize.text.function = function(x){x},
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)
@
\end{center}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp1}
\framesubtitle{IFA model}
Categorical Item Factor Analysis (IFA) model \hyperlink{mimain}{\beamerreturnbutton{Back}}

\bi
\item Threshold model \citep{Muthen1984}
\item Latent propensity to endorse each item $i$ for child $j$ in cohort $c$
\begin{align}\label{eq:propens}
\begin{split}
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{EXT}_{jc} + u_{ijc} \qquad \text{for} \; i=1,\dots, 6 \\
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{INT}_{jc} + u_{ijc} \qquad \text{for} \; i=7,\dots, 11
\end{split}
\end{align}

\item Mean/covariance structure
$$\bm{\mu}_c=\bm{\nu}_c + \bm{\Lambda}_c\bm{\kappa}_c \qquad \text{and} \qquad \Sigma_c = \bm{\Lambda}_c \bm{\Phi}_c \bm{\Lambda}_c^\prime + \bm{\Psi}_c.$$
with $$(\theta^{EXT}_{jc}, \theta^{INT}_{jc}) = \bm{\Theta}_{jc} \sim N(\bm{\kappa}_c, \bm{\Phi}_c) \qquad \text{and} \qquad \bm{u}_{jc} \sim N(\bm{0}, \bm{\Psi}_c)$$

\item Discreteness
$$X_{ijc} = s \qquad \text{if} \; \tau_{s,ic} \leq X^*_{ijc} < \tau_{s+1,ic} \qquad \text{for} \; s=0,1,2$$
with $\tau_{0,ic}=-\infty$ and $\tau_{3,ic}=+\infty$.

\ei
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Key Parameters for Invariance}

\bi
\item Loadings $\lambda_{ci}$
\item Thresholds $\tau_{1,ci}$, $\tau_{2,ci}$
\item Intercepts $\nu_{ci}$
\item Unique variances $\Var(u_{ci})$
\item Latent mean vector $\bm{\kappa}_c$
\item Latent covariance matrix $\bm{\Phi}_c$
\ei

\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp2}
\framesubtitle{Invariance steps \hyperlink{mimain}{\beamerreturnbutton{Back}}}
\Fontvi
\bi
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across cohorts)
  \item $\text{diag}(\bm{\Phi}) = \bm{I}, \quad \bm{\kappa}=\bm{0}, \quad \bm{\nu}=\bm{0}, \quad \text{and} \; \text{diag}(\bm{\Psi})=\bm{I}.$
  \ei

\item \textbf{Threshold invariance} model
  \bi
  \item $\tau_{1,Ai} = \tau_{1,Bi}$ for $i=1,\dots,11$, $\tau_{2,Ai} = \tau_{2,Bi}$ for $i=\{1,2,3,4,7,8,9,10\}$
  \item For all groups: $\text{diag}(\bm{\Phi}) = \bm{I}, \quad \bm{\kappa}=\bm{0}$
  \item For the reference group: $\bm{\nu}_A=\bm{0}, \quad \; \text{diag}(\bm{\Sigma}_A)=\bm{I}$
  \ei
\item \textbf{Threshold and Loading invariance} model
  \bi
  \item $\tau_{1,Ai} = \tau_{1,Bi}$ for $i=1,\dots,11$, $\tau_{2,Ai} = \tau_{2,Bi}$ for $i=\{1,2,3,4,7,8,9,10\}$
  \item $\lambda_{Ai} = \lambda_{Bi}$ for $i=1,\dots,11$
  \item For all groups: $\bm{\kappa}=\bm{0}$
  \item For the reference group: $\bm{\nu}_A=\bm{0}, \quad \; \text{diag}(\bm{\Sigma}_A)=\bm{I}, \quad \text{diag}(\bm{\Phi}_A) = \bm{I}$
  \item Allows comparison of latent factor variances
  \ei
\item \textbf{Threshold, Loading, and Intercept invariance} model
  \bi
  \item $\tau_{1,Ai} = \tau_{1,Bi}$ for $i=1,\dots,11$, $\tau_{2,Ai} = \tau_{2,Bi}$ for $i=\{1,2,3,4,7,8,9,10\}$
  \item $\lambda_{Ai} = \lambda_{Bi}$ for $i=1,\dots,11$
  \item For all groups: $\bm{\nu}=\bm{0}$
  \item For the reference group: $\bm{\kappa}_A=\bm{0}, \quad \; \text{diag}(\bm{\Sigma}_A)=\bm{I}, \quad \text{diag}(\bm{\Phi}_A) = \bm{I}$
  \item Allows comparison of latent factor means and variances
  \ei
\ei

\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Fit comparison}
\begin{table}[ht]
\centering
\small
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[3]],afitab[[4]],afitab[[7]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}",
                    digits = 4
                    ),
             size="\\tiny",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                       \\textbf{Males only} \\\\ ",
                      "\\newline \\textbf{Females only} \\\\ \\midrule ",
                      "\\newline \\textbf{Males and females} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\end{table}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ROBUSTNESS APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[noframenumbering]
\frametitle{Age distribution \hyperlink{mires}{\beamerreturnbutton{Back}}}\label{agehist}
\begin{figure}
\begin{center}
<<AGEHIST, dev='tikz', fig.height=6, fig.width=6, out.width='.6\\linewidth', cache=TRUE>>=
@
\end{center}
\caption{Distribution of age in the BCS and MCS sample}
\end{figure}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{fit4}
\framesubtitle{Fit comparison -- 4 group models \hyperlink{mires}{\beamerreturnbutton{Back}}}

\begin{table}[ht]
\centering
\small
<<FA_FIT_RES4, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[7]],afitab[[8]],afitab[[9]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}C{.05\\textwidth}",
                    digits = 4
                    ),
             size="\\tiny",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                      \\newline \\textbf{(7) not age adjusted} \\\\",
                      "\\newline \\textbf{(8) age adjusted} \\\\\\midrule ",
                      "\\newline \\textbf{(9) 59-62m range} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\end{table}

\end{frame}

\begin{frame}[noframenumbering]
\frametitle{Factor scores vs. Sum scores}
\begin{figure}
<<FACRAW, dev='tikz', fig.height=7, fig.width=7, out.width='.65\\linewidth'>>=
@
\end{figure}
\end{frame}

<<LOESS_REAL, dependson='FA_SCORES_SELECT'>>=
@

\begin{frame}[noframenumbering]
\frametitle{Inequality \hyperlink{meanci}{\beamerreturnbutton{Back}}}\label{loessext}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_EXT, dependson='LOESS_REAL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.ext
@
\end{figure}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{Inequality \hyperlink{meanci}{\beamerreturnbutton{Back}}}\label{loessint}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_INT, dependson='LOESS_REAL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.int
@
\end{figure}
\end{frame}

<<LOESS_INFL, dependson="FA_SCORES_SELECT">>=
@

\begin{frame}[noframenumbering]
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_EXT, dependson='LOESS_INFL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.ext
@
\end{figure}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_INT, dependson='LOESS_INFL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.int
@
\end{figure}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_INFL, dependson='FA_SCORES_SELECT', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

scores2plot$incq_infl <- factor(ntile(scores2plot$faminc_infl, 5))

r.ext.f.c <- lm(EXT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="F"))
r.ext.f.q <- lm(EXT ~ incq_infl*cohort, data=subset(scores2plot, sex=="F"))
r.int.f.c <- lm(INT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="F"))
r.int.f.q <- lm(INT ~ incq_infl*cohort, data=subset(scores2plot, sex=="F"))

r.ext.m.c <- lm(EXT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="M"))
r.ext.m.q <- lm(EXT ~ incq_infl*cohort, data=subset(scores2plot, sex=="M"))
r.int.m.c <- lm(INT ~ faminc_infl*cohort, data=subset(scores2plot, sex=="M"))
r.int.m.q <- lm(INT ~ incq_infl*cohort, data=subset(scores2plot, sex=="M"))

stargazer(r.ext.m.q, r.int.m.q, r.ext.f.q, r.int.f.q,
          column.labels=c("Males","Females"), column.separate=c(2,2),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                             "Income Q2",
                             "Income Q3",
                             "Income Q4",
                             "Income Q5",
                             "MCS dummy",
                             "Income Q2 x MCS",
                             "Income Q3 x MCS",
                             "Income Q4 x MCS",
                             "Income Q5 x MCS"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

\begin{frame}[noframenumbering]
\bibliographystyle{chicago}
\def\bibfont{\tiny}
\bibliography{/Users/giacomomason/Documents/BibTex/Papers-cohorts_nc}
\end{frame}

\end{document}