### 2018-09-28 codes/R/paper/paper_jpube.Rnw
Version with external tex

\documentclass[pdftex,10pt,a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amstext}% for resizing the symbol in math
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[babel]{csquotes}
\usepackage{booktabs}
\usepackage{tocloft}
\usepackage{dsfont}
\usepackage{xcolor}
\usepackage{accents}
\usepackage{rotating}
\usepackage{pdflscape}  % for landscape page
\usepackage{lmodern}  %font
\usepackage{natbib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{bm}
\usepackage{array}
\usepackage{multirow}
\usepackage[multiple]{footmisc}   % for multiple footnotes
\usepackage{textgreek}            % for greek letters in text
\usepackage{ragged2e}             % for table notes justification
\usepackage{setspace} % for spacing
\onehalfspacing

\setlength{\jot}{11pt} % increase vertical spacing in equations
\usepackage{txfonts}
\usepackage[T1]{fontenc}

% define new columns for tables
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\renewcommand{\arraystretch}{1.2}%

% CAPTIONS
\usepackage[textfont={bf,normalsize},labelfont={bf,normalsize}]{caption}     % to allow multiple lines in captions, and customize the look
\captionsetup{justification=raggedright,singlelinecheck=false}

% APPENDICES AND SECTION NUMERATION
\usepackage[nottoc,numbib]{tocbibind} 		% bibliography in TOC
\usepackage[toc,page,title]{appendix}       % appendix ('title' makes sections into 'APPENDIX A')
\newcommand{\tocless}[2]{\bgroup\let\addcontentsline=\nocontentsline#1{#2}\egroup} % for empty section
\numberwithin{equation}{section} %for section eq numeration

% CROSSREFERENCING
\usepackage{hyperref}                            % references
\AtBeginDocument{%
\renewcommand\sectionautorefname{Section}        % custom autoref names
\renewcommand\subsectionautorefname{Section}
}
\newcommand{\aref}[1]{\hyperref[#1]{Appendix~\ref{#1}}} % for appendix sections
\usepackage{cleveref}
%tikz
\usepackage{tikz}
\usetikzlibrary{shapes,intersections}
\usetikzlibrary{positioning,arrows, arrows.meta}

%lists
\usepackage[inline]{enumitem}
\setlist{noitemsep}
\setlist[enumerate]{labelsep=*, leftmargin=*}
\setlist[itemize]{labelsep=*, leftmargin=*}

% page setup (from Racine)
\setlength{\topmargin}{0cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{.0cm}
\setlength{\evensidemargin}{.0cm}
\setlength{\textwidth}{16.5cm}
\setlength{\footskip}{1.5cm}
\setlength{\footnotesep}{0.4cm}

% additional symbols
\newcommand{\me}{\mathrm{e}}
\newcommand{\veps}{\varepsilon}
\newcommand{\E}{\mathbb{E}}
\newcommand{\1}{\mathds{1}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Corr}{\mathrm{Corr}}
\newcommand{\vect}{\mathrm{vec}}
\newcommand{\vecht}{\mathrm{vech}}
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\def\chk{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

% additional commands
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bnm}{\begin{enumerate}}
\newcommand{\enm}{\end{enumerate}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}

\usepackage{authblk}
\title{Inequality in socioemotional skills: a cross-cohort comparison}
\author{Orazio Attanasio \hspace{10pt} Richard Blundell \\  Gabriella Conti \hspace{10pt} Giacomo Mason}
\affil{\emph{Institute for Fiscal Studies} \\ \emph{UCL Department of Economics}}
\date{\today}

\begin{document}

\begin{titlepage}
\maketitle

  \begin{abstract}
    Abstract here
  \end{abstract}

\vspace{30pt}
{\footnotesize 
We thank {\color{orange} participants to the Trans Atlantic Public Economics Seminar }
We are grateful to The Centre for Longitudinal Studies, UCL Institute of Education for the use of the data and to the UK Data Archive and UK Data Service for making them available. However, they bear no responsibility for the analysis or interpretation of the data. 
Giacomo Mason obtained funding thanks to an Economic and Social Research Council graduate scholarship.
}
\end{titlepage}

\input{/Users/giacomomason/Dropbox/cohorts_shared/paper_jpube_1intro.tex}
\input{/Users/giacomomason/Dropbox/cohorts_shared/paper_jpube_2datamethods.tex}
\input{/Users/giacomomason/Dropbox/cohorts_shared/paper_jpube_3resultsconclusion.tex}

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=T, # REPLACE LATER
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_bootstrap.R"))
read_chunk(here("codes/R/5_selectvars.R"))
read_chunk(here("codes/R/6_graphs.R"))
read_chunk(here("codes/R/6_graphs_ineq.R"))
read_chunk(here("codes/R/6_graphs_item.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/99_auxfun.R"))

@

<<PREAMBLE >>=
@

<<LOAD_DATA>>=
@

<<AUXFUN>>=
# load auxiliary functions
@

<<MEANTABLE>>=
# make table with means
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<BCSEFA, dependson='LOAD_DATA'>>=
@

<<BCSEFARES, dependson='BCSEFA'>>=
@

<<MCSEFA, dependson='LOAD_DATA'>>=
@

<<MCSEFARES, dependson='MCSEFA'>>=
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<FA_INV_MAIN>>=
# main models only
@

<<FA_INV_AUX_SYN>>=
# syntax for robustness models
@

<<FA_INV_AUX_GEND>>=
# robustness checks for gender alone
@

<<FA_FIT>>=
# assemble fit indices into matrices
@

<<FA_SCORES_MAIN>>=
# produce scores for main model
@

<<FA_BOOTSCORES_READ>>=
# read bootstrap data
@

<<SELECTDATA, dependson="FA_BOOTSCORES_READ">>=
# define data, variables and formulas for regression and decomposition
@

<<SELECTDATA_OUTC, dependson="SELECTDATA">>=
# load outcomes data
@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<FACINEQ_MAKE>>=
# make factor inequality graphs
@

<<ITEMINEQ_MAKE>>=
# make item-level inequality graphs
@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<REGS_DEP, dependson="SELECTDATA">>=
# run regressions explaining EXT INT
@

<<REGS_OUTC, dependson="SELECTDATA_OUTC">>=
# run regressions explaining adolescent/adult outcomes
@

<<REGS_OUTC_TAB, dependson="REGS_OUTC">>=
# make tables from regressions
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\caption{Quantile differences in scores}\label{tab:iqdiff}
\centering\setstretch{1.2}\footnotesize
<<INEQ_STATS, dependson="SELECTDATA", results='asis'>>=

# income 
tb <- data.frame(
  aggregate(finaldata$faminc10_infl, by=list(finaldata$cohort), FUN = function(x) quantile(x, c(.1, .25, .75, .9), na.rm=T))
)
tb$d9010 <- tb$x[,4]/tb$x[,1]
tb$d7525 <- tb$x[,3]/tb$x[,2]

tb.ext <- data.frame(
  aggregate(finaldata$EXT, by=list(finaldata$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.ext$d9010 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"10%"])
tb.ext$d7525 <- abs(tb.ext$x[,"75%"]-tb.ext$x[,"25%"])
tb.ext$d9050 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"50%"])
tb.ext$d5010 <- abs(tb.ext$x[,"50%"]-tb.ext$x[,"10%"])
rownames(tb.ext) <- tb.ext$Group.1
tb.ext <- t(tb.ext[,-c(1,2)]) # remove aggregated quantiles and transpose
tb.ext <- tb.ext[,c("BCS.M","MCS.M","BCS.F","MCS.F")]

tb.int <- data.frame(
  aggregate(finaldata$INT, by=list(finaldata$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.int$d9010 <- abs(tb.int$x[,"90%"]-tb.int$x[,"10%"])
tb.int$d7525 <- abs(tb.int$x[,"75%"]-tb.int$x[,"25%"])
tb.int$d9050 <- abs(tb.int$x[,"90%"]-tb.int$x[,"50%"])
tb.int$d5010 <- abs(tb.int$x[,"50%"]-tb.int$x[,"10%"])
rownames(tb.int) <- tb.int$Group.1
tb.int <- t(tb.int[,-c(1,2)]) # remove aggregated quantiles and transpose
tb.int <- tb.int[,c("BCS.M","MCS.M","BCS.F","MCS.F")]

# assemble and print
table <- rbind(tb.ext,tb.int)
table <- cbind(data.frame(as.matrix(rep(c("90-10", "75-25", "90-50", "50-10"),2))),
               table)

print(xtable(table,
                    align="L{0cm}L{.15\\textwidth}C{.1\\textwidth}C{.1\\textwidth}C{.1\\textwidth}C{.1\\textwidth}",
                    digits = 2
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,4)),
                              command = c(
                              # header
            " & \\multicolumn{2}{c}{\\textbf{Males}} & \\multicolumn{2}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-3}  \\cmidrule(lr){4-5} 
                       & BCS (1970) & MCS (2000/1) & BCS (1970) & MCS (2000/1)  \\\\
                      \\textbf{Externalising} \\\\ ",
                      "\\textbf{Internalising} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\justify {\scriptsize \emph{Notes}: The table shows differences between quantiles of the distribution of socioemotional skills, by gender and cohort. The distribution is a factor score obtained from the factor model in \autoref{sec:methods}. 
\par}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{landscape}
\begin{table}
\caption{Determinants of Socioemotional Skills}\label{tab:determ}
\centering\footnotesize
<<REGS_ALL, results='asis', dependson="REGS_DEP">>=

# select models
regs <- list(r_ext$BCS.M[["OLS"]],r_ext$MCS.M[["OLS"]],r_ext$BCS.F[["OLS"]],r_ext$MCS.F[["OLS"]],
             r_int$BCS.M[["OLS"]],r_int$MCS.M[["OLS"]],r_int$BCS.F[["OLS"]],r_int$MCS.F[["OLS"]])

# robust SEs
robust_se <- list()
robust_p  <- list()
bootstrap_se  <- list(r_ext$BCS.M[["OLSbootse"]],r_ext$MCS.M[["OLSbootse"]],r_ext$BCS.F[["OLSbootse"]],r_ext$MCS.F[["OLSbootse"]],
                      r_int$BCS.M[["OLSbootse"]],r_int$MCS.M[["OLSbootse"]],r_int$BCS.F[["OLSbootse"]],r_int$MCS.F[["OLSbootse"]])
for (i in 1:length(regs)) {
  cov1              <- sandwich::vcovHC(regs[[i]], type = "HC1")
  ct                <- lmtest::coeftest(regs[[i]], vcov = cov1)
  robust_se[[i]] <- ct[, 2]
  robust_p[[i]] <- ct[, 4]
}

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = bootstrap_se, override.pvalues = robust_p,
               single.row = FALSE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = omitcoefs,
               groups = grouplabs,
               custom.columns = list("p-value" = r_ext$Mp, "p-value" = r_ext$Fp, "p-value" = r_int$Mp, "p-value" = r_int$Fp), custom.col.pos = c(4,6,8,10),
               custom.coef.names = varlabs,
               custom.model.names = c("\\multicolumn{6}{c}{\\textbf{Externalising}} & \\multicolumn{6}{c}{\\textbf{Internalising}} \\\\
                                        \\cmidrule(lr){2-7} \\cmidrule(lr){8-13}
                                        & \\multicolumn{3}{c}{\\textbf{Males}} & \\multicolumn{3}{c}{\\textbf{Females}} &
                                        \\multicolumn{3}{c}{\\textbf{Males}} & \\multicolumn{3}{c}{\\textbf{Females}} \\\\ 
                                        \\cmidrule(lr){2-4} \\cmidrule(lr){5-7} \\cmidrule(lr){8-10} \\cmidrule(lr){11-13} 
                                      & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12) \\\\
                                      & BCS", "MCS", "BCS", "MCS", "BCS", "MCS", "BCS", "MCS"),
               custom.note = "",
               include.rmse = FALSE, include.rsquared = FALSE,
               table = FALSE
               )

@
\justify {\scriptsize \emph{Notes}: The table shows coefficients from linear regressions of children's socioemotional skills at five years of age on family background information. The dependent variable is a factor score obtained from the factor model in \autoref{sec:methods}. Col. (1) and (2) show coefficients and standard errors in parentheses, for male children in the BCS and MCS cohorts separately. The latter are obtained using 1,000 bootstrap repetitions, taking into account the factor estimation stage that precedes the regression. Col. (3) shows the p-value of a test that the coefficient is the same in the two cohorts. Col. (4) to (6) repeat for female children. Col. (7) to (12) repeat for internalising skills. All estimates additionally control for region of birth, mother height, number of previous stillbirths at child's birth, preterm birth, a dummy for missing gestational age, and number of other children in the household at child age 5. See \autoref{tab:harmvar} for a description of the variables used.
\par}
\end{table}
\end{landscape}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<REGS_OUTC_TABPR, dependson="REGS_OUTC">>=
# function to display outcome regression tables
# numoutc = number of outcomes

# EXCLUDING RAW SCORES
printoutctab <- function(tab, numoutc) {
print(xtable(tab,
                    align="R{0cm}L{.3\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}",
                    digits = 3
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = as.list(c(0, 6*seq(1,numoutc-1))),
                              command = c(
                              # header
            " & \\multicolumn{3}{c}{\\textbf{Males}} & \\multicolumn{3}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-4}  \\cmidrule(lr){5-7}
& Mean & \\multicolumn{2}{c}{Coefficients} & Mean & \\multicolumn{2}{c}{Coefficients} \\\\ \\cmidrule(lr){3-4}  \\cmidrule(lr){6-7}
            & (1) & (2) & (3) & (4) & (5) & (6) \\\\",
            rep(" \\midrule", numoutc-1)
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular",
             floating = FALSE
)
}

# INCLUDING RAW SCORES
printoutctab_raw <- function(tab, numoutc) {
print(xtable(tab,
                    align="R{0cm}L{.2\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}",
                    digits = 3
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = as.list(c(0, 8*seq(1,numoutc-1))),
                              command = c(
                              # header
            " & \\multicolumn{4}{c}{\\textbf{Males}} & \\multicolumn{4}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-5}  \\cmidrule(lr){6-9}
& Mean & \\multicolumn{3}{c}{Coefficients} & Mean & \\multicolumn{3}{c}{Coefficients} \\\\ \\cmidrule(lr){3-5}  \\cmidrule(lr){7-9}
            & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) \\\\",
            rep(" \\midrule", numoutc-1)
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular",
             floating = FALSE
)
}
@

\begin{table}
\centering\footnotesize
\caption{Predictors of adolescent outcomes}\label{tab:adol}
<<REGS_OUTC_ADOL, results='asis', dependson="REGS_OUTC">>=

# collect adolescent outcomes
outctabs <- list(bcs_outctab$smktry16, bcs_outctab$bmi16, mcs_outctab$smktry14, mcs_outctab$bmi14)
outctabs <- lapply(outctabs, function(x) x[-4:-5, c(-5,-9)]) # eliminate raw rows and columns
# collapse to matrix
outctabs_mat <- do.call(rbind,lapply(outctabs,matrix,ncol=7,byrow=FALSE))
printoutctab(outctabs_mat, length(outctabs))

@
\justify {\scriptsize \emph{Notes}: The table shows coefficients from linear regressions of cohort members' adolescent outcomes on their externalising and internalising socioemotional skills at five years of age. Col. (1) shows the mean of the outcome for males. Col. (2) regresses the outcome on the scores obtained from the factor model in \autoref{sec:methods}. Col. (3) additionally controls for cognitive ability at age five. This is a simple factor score obtained by aggregating the available cognitive measures. All standard errors in parentheses are obtained using 1,000 bootstrap repetitions, taking into account the factor estimation stage that precedes the regression. Col. (4) to (6) repeat for female cohort members.  All estimates additionally control for region of birth, maternal education (5), maternal employment (5), father occupation (5), maternal background (age, height, nonwhite ethnicity, number of children in HH), pregnancy (firstborn child, number of previous stillbirths, mother smoked in pregnancy, preterm birth, (log) birth weight). See \autoref{tab:harmvar} for a description of the variables used.
\par}
\end{table}

\begin{table}
\centering\footnotesize
\caption{Predictors of adult outcomes -- BCS}\label{tab:bcs_adult}
<<REGS_OUTC_ADULT, results='asis', dependson="REGS_OUTC">>=

# eliminate raw rows and columns
outcs <- c("hnvq34", "emp34", "lgpay34", "emp42", "lgpay42")
outctabs <- list()
for (o in 1:length(outcs)) outctabs[[outcs[o]]] <- bcs_outctab[[outcs[o]]][-4:-5, c(-5,-9)]
# collapse to matrix
outctabs_mat <- do.call(rbind,lapply(outctabs,matrix,ncol=7,byrow=FALSE))
printoutctab(outctabs_mat, length(outcs))

@
\justify {\scriptsize \emph{Notes}: The table shows coefficients from linear regressions of BCS cohort members' adult outcomes on their externalising and internalising socioemotional skills at five years of age. Col. (1) shows the mean of the outcome for males. Col. (2) regresses the outcome on the scores obtained from the factor model in \autoref{sec:methods}. Col. (3) additionally controls for cognitive ability at age five. This is a simple factor score obtained by aggregating the available cognitive measures. All standard errors in parentheses are obtained using 1,000 bootstrap repetitions, taking into account the factor estimation stage that precedes the regression. Col. (4) to (6) repeat for female cohort members.  All estimates additionally control for region of birth, maternal education (5), maternal employment (5), father occupation (5), maternal background (age, height, nonwhite ethnicity, number of children in HH), pregnancy (firstborn child, number of previous stillbirths, mother smoked in pregnancy, preterm birth, (log) birth weight). See \autoref{tab:harmvar} for a description of the variables used.
\par}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
<<FACDENS, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
@
\caption{Distribution of factor scores}\label{fig:dens}
\justify {\scriptsize \emph{Notes}: The figure shows the distribution of the externalising and internalising socioemotional skills scores at age five obtained from the factor model, by gender and cohort. The scores are estimated from parameter estimates in \autoref{tab:measpars}, using an Empirical Bayes Modal approach. Higher scores correspond to \emph{better} skills. The distribution is estimated nonparametrically, using an Epanechnikov kernel. The figure also reports the $p$ value from Kolmogorov-Smirnov tests of equality between the distribution in BCS and MCS.
\par}
\end{figure}

\begin{figure}
<<FACINEQ_PS, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_ps # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Skill inequality by mother's education}\label{fig:ineqps}
\justify {\scriptsize \emph{Notes}: The figure shows unconditional mean values of socioemotional skills scores by gender, cohort, and mother's education at age five. Mother's education is a dummy for whether the mother continued schooling past the minimum leaving age, based on her date of birth. The four panels on top present mean and 95\% confidence intervals. Given that we cannot compare means of skills, all scores are normalised to take value zero for the `Compulsory' category, so that the gradient is emphasised. The bottom two panels present the unconditional distribution of mother's education.
\par}
\end{figure}

\begin{figure}
<<FACINEQ_SM, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_sm # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Skill inequality by mother's pregnancy smoking}\label{fig:ineqsm}
\justify {\scriptsize \emph{Notes}: The figure shows unconditional mean values of socioemotional skills scores by gender, cohort, and mother's pregnancy smoking. Mother's education is a dummy for whether the mother reported smoking during pregnancy. The four panels on top present mean and 95\% confidence intervals. Given that we cannot compare means of skills, all scores are normalised to take value zero for the `Non-smoker' category, so that the gradient is emphasised. The bottom two panels present the unconditional distribution of mother's pregnancy smoking
\par}
\end{figure}

\begin{figure}
<<FACINEQ_FSC, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=10, out.width='\\linewidth'>>=
fi <- fi_fsc # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]],
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Skill inequality by father's occupation}\label{fig:ineqfsc}
\justify {\scriptsize \emph{Notes}: The figure shows unconditional mean values of socioemotional skills scores by gender, cohort, and father's occupation at age five. Father's occupation is based on Registrar General's social class, with classes I to III Non Manual being `White collar' and classes III Manual to V (plus `other') being `Blue collar'. `No father figure' is defined as absence of a male figure living in the household. The four panels on top present mean and 95\% confidence intervals. Given that we cannot compare means of skills, all scores are normalised to take value zero for the `Blue collar' category, so that the gradient is emphasised. The bottom two panels present the unconditional distribution of father's occupation.
\par}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\bibliographystyle{chicago}
\bibliography{/Users/giacomomason/Dropbox/Zotero/cohorts.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\begin{appendices}

% to number tables and figures with A
        \setcounter{table}{0}
        \renewcommand{\thetable}{A\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{A\arabic{figure}}%

% Appendix text
\input{/Users/giacomomason/Dropbox/cohorts_shared/paper_jpube_4appendix.tex}

% Appendix tables
\clearpage
\section{Appendix tables}

% from tex file
\input{/Users/giacomomason/Dropbox/cohorts_shared/paper_jpube_5apptables.tex}

\clearpage
\begin{landscape}
\begin{table}[]
\caption{Item prevalence, by cohort and gender}\label{tab:meantab}
\centering\setstretch{1.2} \footnotesize
<<MEANTAB, dependson="MEANTABLE", results='asis'>>=
require(xtable)
table <- meantab # females
table <- table[,!names(table) %in% c("bcs.text", "bcs.num", "mcs.text", "mcs.num")] # drop if unneeded

  print(xtable(table,
               align="L{0cm}C{.04\\textwidth}C{.05\\textwidth}C{.05\\textwidth}L{.12\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
               digits = 1
  ), 
  include.rownames=F,
  include.colnames=F,
  booktabs = T,
  add.to.row = list(pos = c(list(0)),
                    command = c(
                      # header
                      paste(" & & & &
                            \\multicolumn{6}{c}{\\textbf{Males}} & \\multicolumn{6}{c}{\\textbf{Females}}
                            \\\\ \\cmidrule(lr){5-10} \\cmidrule(lr){11-16} 
                            & & & & 
                            \\multicolumn{3}{c}{\\textbf{BCS}} & \\multicolumn{3}{c}{\\textbf{MCS}}
                            & \\multicolumn{3}{c}{\\textbf{BCS}} & \\multicolumn{3}{c}{\\textbf{MCS}}
                            \\\\ \\cmidrule(lr){5-7} \\cmidrule(lr){8-10} \\cmidrule(lr){11-13} \\cmidrule(lr){14-16} 
                            & & & &
                            Cert. Appl. & Smtm. Appl. & Appl. &  Cert. True & Smwt. True & True &
                            Cert. Appl. & Smtm. Appl. & Appl. &  Cert. True & Smwt. True & True 
                            \\\\
                            Itm. & Factor & Cat. & Title & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) & (\\%) \\\\
                            "
                            , sep="")
                    )),
  sanitize.text.function = function(x){x},
  sanitize.rownames.function = function(x){""}, # to not print row numbers
  tabular.environment = "tabular",
  floating = FALSE
  )

@
\justify {\scriptsize \emph{Notes}: The table shows the prevalence by gender and cohort for each item of our novel subscale. \emph{Itm.} is item number. \emph{Factor} is the latent construct to which the item loads -- EXT is Externalising skills, INT is Internalising skills. \emph{Cat.} is the number of categories in which the item is coded -- 2 denotes a binary item (applies/does not apply) and 3 denotes a 3-category item. \emph{Title} is a short label for the item. \emph{Cert. / Smtm. Appl.} =  Certainly / sometimes applies. \emph{Cert. / Smwt. True} = Certainly / somewhat true.
\par}
\end{table}
\end{landscape}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\caption{Suggested number of factors to retain}\label{tab:numfac}
\centering\setstretch{1.2}\footnotesize
<<NUMFACTAB, dependson='MCSEFARES', results='asis'>>=
# combine results
efatab <- data.frame(cbind(efatab.bcs,efatab.mcs))

efatab %>%
  tibble::rownames_to_column() %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('l','c','c','c','c','c','c'),
        col.names = c("Approach", "All", "Males", "Females", "All", "Males", "Females")) %>%
  add_header_above(c(" " = 1,  "BCS (1970)" = 3, "MCS (2000/1)" = 3)) %>%
  column_spec(2:7, width = "3em")

@
\justify {\scriptsize \emph{Notes}: The table compares the optimal number of factors suggested by different aprroaches for our novel scale; scree test based approaches (optimal coordinates, acceleration factor -- \citealp{Raiche2013}), parallel analysis \citep{Horn1965}, Kaiser's eigenvalue rule \citep{Kaiser1960}, Very Simple Structure (VSS, \citealp{Revelle1979}), Velicer Minimum Average Partial test (MAP, \citealp{Velicer1976}).
\par}
\end{table}

\begin{table}
\caption{Loadings from exploratory factor analysis}\label{tab:load}
\centering\setstretch{1.2}\footnotesize
<<EFALOADTAB, dependson='MCSEFARES', results='asis'>>=
# combine results
library(kableExtra)
loadtab <- cbind(meantab$num, meantab$title, data.frame(cbind(bcsfa$loadings, mcsfa$loadings)))
colnames(loadtab) <- c("item", "title", "bcs1", "bcs2", "mcs1", "mcs2")

loadtab %>%
  mutate_at(vars(matches("cs")), round, 3) %>%
  mutate(
    bcs1 = cell_spec(bcs1, "latex", bold = ifelse(abs(bcs1) > .25, TRUE, FALSE)),
    bcs2 = cell_spec(bcs2, "latex", bold = ifelse(abs(bcs2) > .25, TRUE, FALSE)),
    mcs1 = cell_spec(mcs1, "latex", bold = ifelse(abs(mcs1) > .25, TRUE, FALSE)),
    mcs2 = cell_spec(mcs2, "latex", bold = ifelse(abs(mcs2) > .25, TRUE, FALSE))
  ) %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('c','l','c','c','c'),
        col.names = c("Item", "Title", "Factor 1 (EXT)", "Factor 2 (INT)", "Factor 1 (EXT)", "Factor 2 (INT)")) %>%
  add_header_above(c(" " = 2,  "BCS (1970)" = 2, "MCS (2000/1)" = 2))

@
\justify {\scriptsize \emph{Notes}: The table displays the factor loadings obtained from exploratory factor analysis (EFA) on our novel scale, separately by cohort. The EFA is based on the decomposition of the polychoric correlation matrix, and uses weighted least squares and oblimin rotation.
\par}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{landscape}
\begin{table}[ht]
\caption{Measurement invariance fit comparison}\label{tab:fit}
\centering\setstretch{1.2} \scriptsize
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

# MAKE FIT TABLES
fit_main <- afitab(list(fa.c[[1]], fa.tl[[1]], fa.tli[[1]]), 
       c("Configural", 
         "Threshold + Loading Inv", 
         "Threshold, Loading, + Intercept Inv")
)

fit_agesub <- afitab(list(fa.c[[2]], fa.tl[[2]], fa.tli[[2]]), 
                   c("Configural", 
                     "Threshold + Loading Inv", 
                     "Threshold, Loading, + Intercept Inv")
)
fit_males <- afitab(list(fa.c[[4]], fa.tl[[4]], fa.tli[[4]]), 
                   c("Configural", 
                     "Threshold + Loading Inv", 
                     "Threshold, Loading, + Intercept Inv")
)
fit_females <- afitab(list(fa.c[[5]], fa.tl[[5]], fa.tli[[5]]), 
                   c("Configural", 
                     "Threshold + Loading Inv", 
                     "Threshold, Loading, + Intercept Inv")
)

table <- rbind(fit_main,fit_agesub,fit_males,fit_females)

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.05\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 4
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6,9)),
                              command = c(
                              # header
            "& & \\multicolumn{6}{c}{\\textbf{Absolute fit}} & \\multicolumn{6}{c}{\\textbf{Relative fit}} 
                  \\\\ \\cmidrule(lr){3-8} \\cmidrule(lr){9-14}
            Model & Num. params & $\\chi^2$ & RMSE & SRMR & MFI & CFI & G-hat & $\\chi^2$ $p$ & \\textDelta RMSE & \\textDelta SRMR &  \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ 
& (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12) & (13) \\\\ \\midrule
                       \\textbf{A: Entire sample} \\\\",
                      "\\newline \\textbf{B: 59-61 months sample} \\\\ \\midrule ",
                      "\\newline \\textbf{C: Males only} \\\\ \\midrule ",
                      "\\newline \\textbf{D: Females only} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\justify {\scriptsize \emph{Notes}: The table presents fit indices for models of different invariance levels, following \cite{Wu2016a} Col. (1) displays the number of estimated parameters for each model. Col. (2) and (8) present the value of the $\chi^2$ statistic and the pvalue of the test of equality with respect to the configural model. Col. (3)-(7) and (9)-(13) present alternative fit indices (AFIs), in absolute values and differences from the configural model respectively. RMSE = Root mean squared error of approximation; SRMR = standardised root mean residual; MFI = McDonald non-centrality index; CFI = comparative fit index; G-hat = gamma-hat. Panel A shows results for the whole sample of children in the BCS and MCS cohorts. Panel B is restricted to a subsample of children in the age range of maximum overlap between the two cohorts (59-61 months). Panel C and D are limited to the samples of male and female children respectively.
\par}
\end{table}
\end{landscape}

\clearpage
\begin{table}
\caption{Parameter estimates from factor model with threshold and loading invariance}\label{tab:measpars}
\centering\setstretch{1.2}\footnotesize
\flushleft\textbf{Panel A: Measurement parameters}
<<RES_FA_PAR, dependson="FA_FIT", results='asis'>>=

table <- getmeaspars(finalmod, groups = 4, mode = "tl")

# sequences to add rows
rws <- seq(2, (nrow(table)-1), by = 1) # start from row 2 (exclude model titles)
rwsl <- as.list(rws)

print.xtable(xtable(table,
              align="C{0cm}C{.08\\textwidth}C{.09\\textwidth}C{.07\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
              digits = 3
                    ),
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
                                          " &
                                            & \\textbf{Loadings}
                                            & \\multicolumn{2}{c}{\\textbf{Thresholds}}
                                            & \\multicolumn{3}{c}{\\textbf{Intercepts} (BCS M = 0)}
                                            & \\multicolumn{3}{c}{\\textbf{Variances} (BCS M = 1)}
                                            \\\\ \\cmidrule(lr){3-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-8} \\cmidrule(lr){9-11}

                                            &
                                            & $\\lambda$
                                            & $\\tau_1$ & $\\tau_2$
                                            & \\multicolumn{3}{c}{$\\nu$}
                                            & \\multicolumn{3}{c}{$\\text{diag}(\\Psi)$} \\\\

                                            Item
                                            & Factor
                                            & All
                                            & All & All
                                            & BCS F & MCS M & MCS F
                                            & BCS F & MCS M & MCS F
                                            \\\\ "
                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
@
\vspace{3pt}
\flushleft\textbf{Panel B: Latent variable parameters}
<<RES_FA_LV, dependson="FA_FIT", results='asis'>>=
require(xtable)

lvtab <- getlvpars(finalmod, groups = 4)
table <- rbind(lvtab$Males,lvtab$Females)
table <- table[,c("measure", "mean_BCS", "mean_MCS", 
                  "covext_BCS", "covint_BCS", "covext_MCS", "covint_MCS", 
                  "corr_BCS", "corr_MCS")] # rearrange

print.xtable(xtable(table,
                     align="C{0cm}L{.13\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 3
                    ),
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,2,4)),
                              command = c(
                              # header
                      "\\toprule
                      & \\multicolumn{2}{c}{\\textbf{Mean}}
                      & \\multicolumn{4}{c}{\\textbf{Covariance}}
                      & \\multicolumn{2}{c}{Correlation} \\\\
                      \\cmidrule(lr){2-3} \\cmidrule(lr){4-7} \\cmidrule(lr){8-9}

                      & \\multicolumn{2}{c}{$\\bm{\\kappa}$} & \\multicolumn{4}{c}{$\\bm{\\Phi}$} & & \\\\

                      & BCS & MCS & \\multicolumn{2}{c}{BCS} & \\multicolumn{2}{c}{MCS} & BCS & MCS \\\\

                      \\textbf{Males} \\\\",
                      "\\textbf{Females} \\\\",
                      "\\bottomrule"
                                    )),
             hline.after = NULL,  # remove default booktabs lines
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\justify {\scriptsize \emph{Notes}: The table presents estimates for the factor model with loadings $\lambda$ and thresholds $\tau$ restricted to be equal across cohorts. Panel A shows estimates of the measurement parameters. Loadings and thresholds are the same across all cohorts. Intercepts are restricted to zero in the reference group, i.e. males in BCS (not shown). Variances of the error terms are restricted to one in the reference group, i.e. males in BCS (not shown), and for the items that only have two categories (5, 6, 11). Panel B shows estimates of the latent variable parameters. Means are restricted to zero in all cohort-gender groups, while variances are restricted to one only in the reference group, i.e. males in BCS.
\par}
\end{table}

\begin{table}
\centering\scriptsize
\caption{Predictors of adolescent outcomes, with sum scores}\label{tab:adol_app}
\resizebox{15cm}{!}{
<<REGS_OUTC_ADOL_APP, results='asis', dependson="REGS_OUTC">>=

# collect adolescent outcomes
outctabs <- list(bcs_outctab$smktry16, bcs_outctab$bmi16, mcs_outctab$smktry14, mcs_outctab$bmi14)
# collapse to matrix
outctabs_mat <- do.call(rbind,lapply(outctabs,matrix,ncol=9,byrow=FALSE))
printoutctab_raw(outctabs_mat, length(outctabs))

@
}
\justify {\scriptsize \emph{Notes}: The table shows coefficients from linear regressions of cohort members' adolescent and adult outcomes on their externalising and internalising socioemotional skills at five years of age. Col. (1) shows the mean of the outcome for males. Col. (2) regresses the outcome on the scores obtained from the factor model in \autoref{sec:methods}. Col. (3) additionally controls for cognitive ability at age five. This is a simple factor score obtained by aggregating the available cognitive measures. Col. (4) uses sum scores (see \autoref{fig:sumscores}) instead of factor scores. All standard errors in parentheses are obtained using 1,000 bootstrap repetitions, taking into account the factor estimation stage that precedes the regression. Col. (5) to (8) repeat for female cohort members.  All estimates additionally control for region of birth, maternal education (5), maternal employment (5), father occupation (5), maternal background (age, height, nonwhite ethnicity, number of children in HH), pregnancy (firstborn child, number of previous stillbirths, mother smoked in pregnancy, preterm birth, (log) birth weight). See \autoref{tab:harmvar} for a description of the variables used.
\par}
\end{table}

\begin{table}
\centering\footnotesize
\caption{Predictors of adult behaviours, BCS}\label{tab:bcs_adultapp}
<<REGS_OUTC_BCS_APP, results='asis', dependson="REGS_OUTC">>=

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c("smoke42", "bmi42")],matrix,ncol=9,byrow=FALSE))
printoutctab_raw(bcs_outctab_mat, 2)

@
\justify {\scriptsize \emph{Notes}: The table shows coefficients from linear regressions of cohort members' adolescent and adult outcomes on their externalising and internalising socioemotional skills at five years of age. Col. (1) shows the mean of the outcome for males. Col. (2) regresses the outcome on the scores obtained from the factor model in \autoref{sec:methods}. Col. (3) additionally controls for cognitive ability at age five. This is a simple factor score obtained by aggregating the available cognitive measures. Col. (4) uses sum scores (see \autoref{fig:sumscores}) instead of factor scores. All standard errors in parentheses are obtained using 1,000 bootstrap repetitions, taking into account the factor estimation stage that precedes the regression. Col. (5) to (8) repeat for female cohort members.  All estimates additionally control for region of birth, maternal education (5), maternal employment (5), father occupation (5), maternal background (age, height, nonwhite ethnicity, number of children in HH), pregnancy (firstborn child, number of previous stillbirths, mother smoked in pregnancy, preterm birth, (log) birth weight). See \autoref{tab:harmvar} for a description of the variables used.
\par}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Appendix figures
\clearpage
\section{Appendix figures}

\begin{figure}[ht]
<<RAWHIST, dependson="SELECTDATA", dev='tikz', fig.height=9, fig.width=10, out.width='\\linewidth'>>=
@
\caption{Distribution of sum scores}\label{fig:sumscores}
\justify {\scriptsize \emph{Notes}: The figure shows the distribution of the externalising and internalising sum scores at age five, by gender and cohort. The scores are obtained by assigning 0, 1, or 2 points for each item in the scale in \autoref{tab:scalecomp}. Zero points are assigned for `Certainly Applies / True' responses, one point for `Sometimes applies / somewhat true', and two points for `Doesn't apply'. Only 0 or 1 points are assigned for items that are coded as having two categories (5,6, and 11). Higher scores correspond to \emph{better} skills.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_PS, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=11, out.width='\\linewidth'>>=
itempl_ps
@
\caption{Item-level inequality by mother's education}\label{fig:ineqips}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between the prevalence of each item in our scale in children of educated vs uneducated mothers, by cohort and gender. All items that have three categories in the scale have been dichotomised. For example, if the prevalence of the `Restless' behaviours among children of mothers with compulsory schooling in the BCS cohort is 7.5\%, and 5\% among mothers with post-compulsory schooling, the ratio will be 1.5. The error bars represent 95\% confidence intervals.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_SM, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=11, out.width='\\linewidth'>>=
itempl_sm
@
\caption{Item-level inequality by mother's pregnancy smoking}\label{fig:ineqism}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between the prevalence of each item in our scale in children of mothers who smoked in pregnancy vs non-smokers, by cohort and gender. All items that have three categories in the scale have been dichotomised. For example, if the prevalence of the `Restless' behaviours among children of smoker mothers in the BCS cohort is 7.5\%, and 5\% among non-smoker mothers, the ratio will be 1.5. The error bars represent 95\% confidence intervals.
\par}
\end{figure}

\begin{figure}
<<ITEMINEQ_FSC, dependson="SELECTDATA", dev='tikz', fig.height=10, fig.width=11, out.width='\\linewidth'>>=
itempl_fsc
@
\caption{Item-level inequality by father's occupation}\label{fig:ineqifsc}
\justify {\scriptsize \emph{Notes}: The graph displays the ratio between the prevalence of each item in our scale in children of white collar vs blue collar fathers, by cohort and gender. All items that have three categories in the scale have been dichotomised. For example, if the prevalence of the `Restless' behaviours among children of blue collar fathers in the BCS cohort is 7.5\%, and 5\% among white collar fathers, the ratio will be 1.5. The error bars represent 95\% confidence intervals.
\par}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{appendices}

\end{document}