### 2020-02-11 codes/R/paper/text/paper_jpube_4appendix.tex

\section{Deriving a common scale of socio-emotional skills}\label{asec:scales}

In the BCS, maternal reports on child socio-emotional skills are measured using the Rutter A Scale \citep{Rutter1970} -- see Panel A of \autoref{tab:scales}. The Rutter items are rated on three levels: `Does not apply', `Somewhat applies', `Certainly applies'. Since they all indicate negative behaviours, we recode all of them in reverse, i.e. `Certainly applies' = 0, `Somewhat applies' = 1, `Does not apply' = 2. We augment the 19-item Rutter Scale with three additional parent-reported questions from the parental questionnaire, items A, B, and D. These are rated on 4 levels: `Never in the last 12 months', `less than once a month', `at least once a month', `at least once a week'; we recode them into binary indicators, with `Never' and `Less than once a month' to 1 and zero otherwise. To increase comparability between the two scales, we also merge together two pairs of Rutter items: 4 and 19 (to mirror SDQ item 12 ``Often fights with other children or bullies them"), and A and B (to mirror SDQ item 3 ``Often complains of head-aches, stomach-ache or sickness"); we assign the lowest category among the two original items to the newly obtained item. We also recode the three-category Rutter items 5 and 14 to binary to mimic the split in the MCS, where they are worded positively.

In the MCS, we use the 25-item Strengths and Difficulties Questionnaire \citep{Goodman1997} -- see Panel B of \autoref{tab:scales}. All items are recorded on a 4-point scale: `Not true', `Somewhat true', `Certainly true', `Can't say'. We set the latter option to missing and recode the rest so that a greater value represents a higher level of skills, as for the BCS items, i.e. `Certainly true' = 0, `Somewhat true' = 1, `Not true' = 2 for the negatively-worded items (and the opposite for the positively-worded ones). For comparability with the BCS Rutter scale, we dichotomise items 3 and 5, and dichotomise and invert items 7 and 14.

\section{Robustness of exploratory analysis}\label{asec:fullefa}

In this section, we repeat the exploratory analysis step in \autoref{sec:ea} for the full set of Rutter and SDQ items. This is to show that the factor structure emerging from the exploratory analysis of the 11-item subscale is consistent with what would emerge considering the original scales in their entirety. Again, we proceed by first assessing the optimal number of factors, and then examining the loadings obtained from exploratory factor analysis.

Results for the optimal number of factors as indicated by different approaches are in \autoref{tab:numfac_all}. Similarly to the 11-item subscale, there is not much agreement between methods. Since the purpose of this section is to assess the robustness of the 11-item subscale, we adopt a conservative approach by estimating EFA models with the largest number of factors suggested, i.e. five. In this way, we allow for richer factor solutions, that have more power to disprove our simpler two-factor solution for the novel subscale.

\autoref{tab:load_bcsall} presents factor loadings for the Rutter scale in BCS, with the addition of the "headaches/stomachaches" and "tantrums" items (see Appendix \autoref{asec:scales}). The split between externalising and internalising items that we recover in the 11-item scale is almost entirely preserved in the full scale, as seen by items loading on factors 1 and 2. The only exception is the "headaches/stomachaches" item, which seems to load on a separate factor. We then carry out robustness checks for the measurement invariance analysis excluding this item in Appendix \autoref{asec:no511}.

The same analysis is repeated for the SDQ scale in MCS in \autoref{tab:load_mcsall}. An internalising, emotional dimension (factor 3) emerges neatly, and coherently with the analysis on our subscale. The externalising items from our subscales are split across two dimensions in this full-scale EFA: one more related to hyperactivity (factor 2) and one to conduct problems (factor 4). This is consistent with the original structure of the SDQ \citep{Goodman1997}.

\section{Robustness of item choice}\label{asec:no511}

In deriving our novel 11-item scale, we construct two items for the BCS cohort based on questions that are not in the original Rutter scale -- namely those concerning "headaches/stomachaches" and "tantrums" (see Appendix \autoref{asec:scales} above for details). Concerns might arise that introducing these items might somewhat invalidate our main conclusions, rather than provide additional informational content on children's externalising and internalising behaviours and symptoms.

In fact, exploratory factor analysis (on both the full Rutter and SDQ scales and on the 11-item subscale) shows that these items, numbered 5 and 11, perform poorly and exhibit relatively low factor loadings. As a robustness check, we replicate the main results of the paper by excluding them from the subscale.

Panel C of \autoref{tab:fit} shows that the measurement invariance analysis yields the same qualitative results once these two items are included. \autoref{fig:scatter_no511} shows a scatter plot of the factor scores obtained from the factor model with and without items 5 and 11. They exhibit very high correlation, thus indicating that our results in \autoref{sec:results} would not substantially change if we omitted the two items with the least informational content.